---
title: "Abdominal Imaging and HCC Risk"
author: "Shelley Facente and Steph Holm"
date: "March 4, 2020"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(pacman)
p_load(tidyverse, readxl, skimr, knitr, xslx, lubridate, MASS, ggplot2, ggpubr, data.table, mice, ltmle, SuperLearner, arm, gam, randomForest, earth, Amelia)
#library(Hmisc)

knitr::opts_chunk$set(include = FALSE, cache = TRUE)

# First, set up data loading functions
StephandShelleyDataLoad <- function(filename) 
{
  if(Sys.info()['sysname']=="Darwin"){
    
     read_excel(as.character(filename))}
  else{
    windowsfilepath <- paste("~/GitHub/stephandshelley/", filename, sep = "")
     read_excel(as.character(windowsfilepath))
  }
}

StephandShelleyDataLoad2 <- function(filename) 
{
  if(Sys.info()['sysname']=="Darwin"){
    
     load(as.character(filename))}
  else{
    windowsfilepath <- paste("~/GitHub/stephandshelley/", filename, sep = "")
     load(as.character(windowsfilepath))
  }
}

```

```{r Data_Cleaning, include = FALSE}

# HCVData <- StephandShelleyDataLoad("Deidentified Chronic HCV Dataset 11-21-19.xlsx")
# 
# #first just making variables their proper type, and renaming so there's no spaces
# HCVData$`Deidentifier  11-21-19` <-  as.factor(HCVData$`Deidentifier  11-21-19`)
# colnames(HCVData)[colnames(HCVData) == 'Deidentifier  11-21-19'] <- 'PtID'
# 
# HCVData$'Hep Visit?'<- as.factor(HCVData$'Hep Visit?')
# colnames(HCVData)[colnames(HCVData) == 'Hep Visit?'] <- 'EverHepVisit'
# 
# HCVData$'Ever Chronic HCV?' <- as.factor(HCVData$'Ever Chronic HCV?')
# colnames(HCVData)[colnames(HCVData) == 'Ever Chronic HCV?'] <- 'EverChronicHCV'
# 
# HCVData$'Spontaneous Clearance?' <- as.factor(HCVData$'Spontaneous Clearance?')
# colnames(HCVData)[colnames(HCVData) == 'Spontaneous Clearance?'] <- 'SpontClearance'
# 
# HCVData$'Ever Ab positive HCV?' <- as.factor(HCVData$'Ever Ab positive HCV?')
# colnames(HCVData)[colnames(HCVData) == 'Ever Ab positive HCV?'] <- 'EverAbPosHCV'
# 
# HCVData$'Ever Treated?'<- as.factor(HCVData$'Ever Treated?')
# colnames(HCVData)[colnames(HCVData) == 'Ever Treated?'] <- 'EverTreated'
# 
# HCVData$'Evidence of SVR?'<- as.factor(HCVData$'Evidence of SVR?')
# colnames(HCVData)[colnames(HCVData) == 'Evidence of SVR?'] <- 'SVRAchieved'
# 
# HCVData$'Treatment Definition' <- as.factor(HCVData$'Treatment Definition')
# colnames(HCVData)[colnames(HCVData) == 'Treatment Definition'] <- 'TreatmentDefinition'
# 
# HCVData$'DX_TYPE OR TEST_TYPE' <- as.factor(HCVData$'DX_TYPE OR TEST_TYPE')
# colnames(HCVData)[colnames(HCVData) == 'DX_TYPE OR TEST_TYPE'] <- 'inclusion.source'
# 
# HCVData$'DX_NAME OR TEST RESULT'<- as.factor(HCVData$'DX_NAME OR TEST RESULT')
# colnames(HCVData)[colnames(HCVData) == 'DX_NAME OR TEST RESULT'] <- 'inclusion.reason'
# 
# #Didn't do anything with the column 'ICD10 OR RESULT_DATE' because it's a mix of formats/info
# 
# colnames(HCVData)[colnames(HCVData) == 'SPECIMN TAKEN DATE'] <- 'date.first.pos.HCV.test'
# 
# HCVData$RESULT <- as.factor(HCVData$RESULT)
# colnames(HCVData)[colnames(HCVData) == 'RESULT'] <- 'HCV.test.result'
# 
# HCVData$'RNA EVER TESTED' <- as.factor(HCVData$'RNA EVER TESTED')
# colnames(HCVData)[colnames(HCVData) == 'RNA EVER TESTED'] <- 'RNAeverTested'
# 
# colnames(HCVData)[colnames(HCVData) == 'FIRST POS (Or if no POS, "SEE TEXT") RNA TEST DATE'] <- 'Date.First.Pos.RNA'
# 
# HCVData$'FIRST POS (Or if no POS, "SEE TEXT") RNA TEST RESULT' <- as.factor(HCVData$'FIRST POS (Or if no POS, "SEE TEXT") RNA TEST RESULT')
# colnames(HCVData)[colnames(HCVData) == 'FIRST POS (Or if no POS, "SEE TEXT") RNA TEST RESULT'] <- 'Result.First.Pos.RNA'
# 
# #Added this to address the "See Text" weirdness; I think "See Text" was inserted here when the test was undetectable or lab error, etc. ##### MAY WANT TO REMOVE IF NUMERIC PREFERRED
# ##Note from Steph- it's not undetectable/lab error, it's just that for a number of years, they used this as a way to 'hide' results so that people in someone's chart inadvertantly didn't see sensitive test results. It was sort of a weird thing and they eventually dropped it. But that's why it's there. Right now though this is erroring over the RNA.interpretation column so I've commented it out.
# #HCVData$Result.First.Pos.RNA <- ifelse(HCVData$Result.First.Pos.RNA == 'See Text', HCVData$RNA.interpretation, HCVData$Result.First.Pos.RNA)
# 
# colnames(HCVData)[colnames(HCVData) == 'MOST RECENT RNA POS'] <- 'Most.Recent.RNA.Pos'
# 
# colnames(HCVData)[colnames(HCVData) == 'MOST RECENT RNA NEG'] <- 'Most.Recent.RNA.Neg'
# 
# #This column is in the midst of the RNA testing section but is unlabeled. It's a date.
# colnames(HCVData)[colnames(HCVData) == '..19'] <- 'UnknownColumn'
# 
# colnames(HCVData)[colnames(HCVData) == 'MOST RECENT RNA "SEE TEXT"'] <- 'Most.Recent.RNA.SeeText'
# 
# colnames(HCVData)[colnames(HCVData) == 'RNA interpretation'] <- 'RNA.interpretation'
# 
# colnames(HCVData)[colnames(HCVData) == 'other notes on diagnosis'] <- 'Notes'
# 
# HCVData$'Where getting care' <- as.factor(HCVData$'Where getting care')
# colnames(HCVData)[colnames(HCVData) == 'Where getting care'] <- 'Care.Location'
# 
# #This errors because there are a number of non-dates in this column. Should discuss with Rena if using.
# #HCVData$'HEP C GENOTYPE DATE' <- as.POSIXct(HCVData$'HEP C GENOTYPE DATE')
# #colnames(HCVData)[colnames(HCVData) == 'HEP C GENOTYPE DATE'] <- 'HCV.Genotype.Date'
# 
# HCVData$'HEP C GENOTYPE RESULT' <- as.factor(HCVData$'HEP C GENOTYPE RESULT')
# colnames(HCVData)[colnames(HCVData) == 'HEP C GENOTYPE RESULT'] <- 'HCV.Genotype.Result'
# 
# colnames(HCVData)[colnames(HCVData) == 'ALPHAFETOPROTEIN_RESULT_DATE'] <- 'AFP.Date'
# 
# colnames(HCVData)[colnames(HCVData) == 'ALPHAFETOPROTEIN_RESULT'] <- 'AFP.Result'
# 
# colnames(HCVData)[colnames(HCVData) == 'FIRST_HEPATOLOGY_APPT'] <- 'First.Hep.Appt'
# 
# colnames(HCVData)[colnames(HCVData) == 'MOST_RECENT_HEPATOLOGY_APPT'] <- 'Most.Recent.Hep.Appt'
# 
# colnames(HCVData)[colnames(HCVData) == 'PLATLET_RESULT_DATE'] <- 'Lowest.Platelet.Date'
# 
# colnames(HCVData)[colnames(HCVData) == 'LOWEST_PLATELETS_RESULT'] <- 'Lowest.Platelet.Result'
# 
# colnames(HCVData)[colnames(HCVData) == 'HEPA_VACCINE_DATE'] <- 'HAV.Vax.Date'
# 
# colnames(HCVData)[colnames(HCVData) == 'HEPB_VACCINE_DATE'] <- 'HBV.Vax.Date'
# 
# HCVData$'Ever Had a Primary Care Visit' <- as.factor(HCVData$'Ever Had a Primary Care Visit')
# colnames(HCVData)[colnames(HCVData) == 'Ever Had a Primary Care Visit'] <- 'EverPrimaryCareVisit'
# 
# HCVData$'DEPT_NAME' <- as.factor(HCVData$'DEPT_NAME')
# 
# HCVData$'CURRENT_PCP' <- as.factor(HCVData$'CURRENT_PCP')
# 
# HCVData$'CURRENT_PCP_DEPT' <- as.factor(HCVData$'CURRENT_PCP_DEPT')
# 
# colnames(HCVData)[colnames(HCVData) == 'HBsAb_RESULT_DATE'] <- 'HBsAG.Date'
# 
# HCVData$'HBsAb_ORD_VALUE' <- as.factor(HCVData$'HBsAb_ORD_VALUE')
# colnames(HCVData)[colnames(HCVData) == 'HBsAb_ORD_VALUE'] <- 'HBsAb.Result'
# 
# colnames(HCVData)[colnames(HCVData) == 'HIV_RESULT_DATE'] <- 'HIV.Date'
# 
# HCVData$'ORD_VALUE' <- as.factor(HCVData$'ORD_VALUE')
# colnames(HCVData)[colnames(HCVData) == 'ORD_VALUE'] <- 'HIV.Result'
# 
# HCVData$'HepA_IgG_Result_Date' <- as.factor(HCVData$'HepA_IgG_Result_Date')
# colnames(HCVData)[colnames(HCVData) == 'HepA_IgG_Result_Date'] <- 'HepA_IgG.Date'
# 
# HCVData$'HepA_IgG' <- as.factor(HCVData$'HepA_IgG')
# colnames(HCVData)[colnames(HCVData) == 'HepA_IgG'] <- 'HepA_IgG.Result'
# 
# HCVData$'HBsAg_RESULT_DATE' <- as.factor(HCVData$'HBsAg_RESULT_DATE')
# colnames(HCVData)[colnames(HCVData) == 'HBsAg_RESULT_DATE'] <- 'HBsAg_Date'
# 
# HCVData$'HBsAg_Value' <- as.factor(HCVData$'HBsAg_Value')
# 
# HCVData$'Med_1' <- as.factor(HCVData$'Med_1')
# 
# HCVData$'Med_2' <- as.factor(HCVData$'Med_2')
# 
# HCVData$'HEPB_CORE_IgM_RESULT' <- as.factor(HCVData$'HEPB_CORE_IgM_RESULT')
# 
# HCVData$'HEPB_CORE_Total_RESULT' <- as.factor(HCVData$'HEPB_CORE_Total_RESULT')
# 
# HCVData$DX_TYPE <- as.factor(HCVData$DX_TYPE)
# colnames(HCVData)[colnames(HCVData) == 'DX_TYPE'] <- 'Cancer.Diagnosis.Source'
# 
# HCVData$'DX_NAME..56' <- as.factor(HCVData$'DX_NAME..56')
# colnames(HCVData)[colnames(HCVData) == 'DX_NAME..56'] <- 'Cancer.Diagnosis.Name'
# 
# HCVData$ICD10 <- as.factor(HCVData$ICD10)
# colnames(HCVData)[colnames(HCVData) == 'ICD10'] <- 'Cancer.ICD10'
# 
# HCVData$FIRST_DX_DATE <- as.numeric(HCVData$FIRST_DX_DATE)
# HCVData$FIRST_DX_DATE <- as.Date(HCVData$FIRST_DX_DATE, origin = "1899-12-30")
# colnames(HCVData)[colnames(HCVData) == 'FIRST_DX_DATE'] <- 'Cancer.Diagnosis.Date'
# 
# HCVData$RACE <- as.factor(HCVData$RACE)
# colnames(HCVData)[colnames(HCVData) == 'RACE'] <- 'Race'
# 
# HCVData$ETHNICTY <- as.factor(HCVData$ETHNICTY)
# colnames(HCVData)[colnames(HCVData) == 'ETHNICTY'] <- 'Ethnicity'
# 
# #Appears to only be male/female, I suspect really sex :(
# HCVData$SEX <- as.factor(HCVData$SEX)
# colnames(HCVData)[colnames(HCVData) == 'SEX'] <- 'Sex'
# 
# HCVData$LANGUAGE <- as.factor(HCVData$LANGUAGE)
# colnames(HCVData)[colnames(HCVData) == 'LANGUAGE'] <- 'Language'
# 
# colnames(HCVData)[colnames(HCVData) == 'Outpatient Completed'] <- 'Num.Outpatient.Visits'
# 
# colnames(HCVData)[colnames(HCVData) == 'Outpatient Canceled'] <- 'Num.Outpatient.Cancellations'
# 
# colnames(HCVData)[colnames(HCVData) == 'Outpatient No Show'] <- 'Num.Outpatient.NoShows'
# 
# colnames(HCVData)[colnames(HCVData) == 'Opiod'] <- 'Opioid.Hx'
# 
# colnames(HCVData)[colnames(HCVData) == 'ED'] <- 'ED.Visits'
# 
# colnames(HCVData)[colnames(HCVData) == 'Hosp Admission'] <- 'Hospital.Admissions'
# 
# ##removing extra variables with the same column names, unneeded for our analysis as far as I can tell
# HCVData <- subset(HCVData, select = -c(DX_NAME..71,
#            CURRENT_ICD10_LIST..72,
#            DX_NAME..74,
#            CURRENT_ICD10_LIST..75,
#            DX_NAME..77,
#            CURRENT_ICD10_LIST..78,
#            DX_NAME..80,
#            CURRENT_ICD10_LIST..81,
#            DX_NAME..83,
#            CURRENT_ICD10_LIST..84,
#            DX_NAME..86,
#            CURRENT_ICD10_LIST..87,
#            DX_NAME..89,
#            CURRENT_ICD10_LIST..90,
#            DX_NAME..92,
#            CURRENT_ICD10_LIST..93,
#            DX_NAME..95,
#            CURRENT_ICD10_LIST..96,
#            DX_NAME..98,
#            CURRENT_ICD10_LIST..99,
#            DX_NAME..101,
#            CURRENT_ICD10_LIST..102,
#            CURRENT_ICD10_LIST..104,
#            DX_NAME..105,
#            CURRENT_ICD10_LIST..107,
#            DX_NAME..108,
#            DX_NAME..110,
#            CURRENT_ICD10_LIST..111,
#            CURRENT_ICD10_LIST..113,
#            DX_NAME..114,
#            CURRENT_ICD10_LIST..116,
#            DX_NAME..117,
#            DX_NAME..119,
#            CURRENT_ICD10_LIST..120,
#            DX_NAME..122,
#            CURRENT_ICD10_LIST..123,
#            DX_NAME..125,
#            CURRENT_ICD10_LIST..126))
# 
# colnames(HCVData)[colnames(HCVData) == 'METASTATIC Cancer'] <- 'Metastatic.Cancer'
# 
# colnames(HCVData)[colnames(HCVData) == 'Nonmetastatic Cancer'] <- 'NON-Metastatic.Cancer'
# 
# colnames(HCVData)[colnames(HCVData) == 'Alcohol Abuse'] <- 'Alcohol.Abuse'
# 
# colnames(HCVData)[colnames(HCVData) == 'Drug Abuse'] <- 'Drug.Abuse'
# 
# colnames(HCVData)[colnames(HCVData) == 'Othe Neurological Disorder'] <- 'Other.Neuro.Disorder'
# 
# colnames(HCVData)[colnames(HCVData) == 'CHRONIC PULMONARY DISORDER'] <- 'CPD'
# 
# colnames(HCVData)[colnames(HCVData) == 'Renal Failure'] <- 'Renal.Failure'
# 
# colnames(HCVData)[colnames(HCVData) == 'Liver Disease'] <- 'Liver.Disease'
# 
# HCVData$PAYOR_NAME <- as.factor(HCVData$PAYOR_NAME)
# colnames(HCVData)[colnames(HCVData) == 'PAYOR_NAME'] <- 'PAYOR'
# 
# HCVData <- HCVData %>% mutate(PayorType = ifelse(str_detect(HCVData$PAYOR, "MCAL"), "MediCal", "Commercial"),
#                               PayorType = ifelse(str_detect(HCVData$PAYOR, "M-CAL"), "MediCal", PayorType),
#                               PayorType = ifelse(str_detect(HCVData$PAYOR, "MEDI-CAL"), "MediCal", PayorType),
#                               PayorType = ifelse(str_detect(HCVData$PAYOR, "SENIOR"), "Medicare", PayorType),
#                               PayorType = ifelse(str_detect(HCVData$PAYOR, "SNR"), "Medicare", PayorType),
#                               PayorType = ifelse(str_detect(HCVData$PAYOR, "MEDICARE"), "Medicare", PayorType),
#                               PayorType = ifelse(PayorType=="MediCal", "MediCal", "NotMediCal"))
# 
# HCVData$'FIRST PC APPT DEPT' <- as.factor(HCVData$'FIRST PC APPT DEPT')
# colnames(HCVData)[colnames(HCVData) == 'FIRST PC APPT DEPT'] <- 'FIRST_PC_APPT_DEPT'
# 
# HCVData$'MOST RECENT PC APPT DEPT' <- as.factor(HCVData$'MOST RECENT PC APPT DEPT')
# colnames(HCVData)[colnames(HCVData) == 'MOST RECENT PC APPT DEPT'] <- 'MOST_RECENT_PC_APPT_DEPT'
# 
# HCVData$VISIT_COUNT_2011 <-rowSums(cbind(HCVData$PC_2011_VISIT_COUNT, HCVData$HEP_2011_VISIT_COUNT), na.rm=TRUE)
# HCVData$VISIT_COUNT_2012 <-rowSums(cbind(HCVData$PC_2012_VISIT_COUNT, HCVData$HEP_2012_VISIT_COUNT), na.rm=TRUE)
# HCVData$VISIT_COUNT_2013 <-rowSums(cbind(HCVData$PC_2013_VISIT_COUNT, HCVData$HEP_2013_VISIT_COUNT), na.rm=TRUE)
# HCVData$VISIT_COUNT_2014 <-rowSums(cbind(HCVData$PC_2014_VISIT_COUNT, HCVData$HEP_2014_VISIT_COUNT), na.rm=TRUE)
# HCVData$VISIT_COUNT_2015 <-rowSums(cbind(HCVData$PC_2015_VISIT_COUNT, HCVData$HEP_2015_VISIT_COUNT), na.rm=TRUE)
# HCVData$VISIT_COUNT_2016 <-rowSums(cbind(HCVData$PC_2016_VISIT_COUNT, HCVData$HEP_2016_VISIT_COUNT), na.rm=TRUE)
# HCVData$VISIT_COUNT_2017 <-rowSums(cbind(HCVData$PC_2017_VISIT_COUNT, HCVData$HEP_2017_VISIT_COUNT), na.rm=TRUE)
# HCVData$VISIT_COUNT_2018 <-rowSums(cbind(HCVData$PC_2018_VISIT_COUNT, HCVData$HEP_2018_VISIT_COUNT), na.rm=TRUE)
# HCVData$VISIT_COUNT_2019 <-rowSums(cbind(HCVData$PC_2019_VISIT_COUNT, HCVData$HEP_2019_VISIT_COUNT), na.rm=TRUE)
# 
# HCVData$TotalVisits<- HCVData$VISIT_COUNT_2011 + HCVData$VISIT_COUNT_2012 + HCVData$VISIT_COUNT_2013 + HCVData$VISIT_COUNT_2014 + HCVData$VISIT_COUNT_2015 + HCVData$VISIT_COUNT_2016 + HCVData$VISIT_COUNT_2017 + HCVData$VISIT_COUNT_2018
# 
# HCVData$TotalVisits15to18 <- HCVData$VISIT_COUNT_2015 + HCVData$VISIT_COUNT_2016 + HCVData$VISIT_COUNT_2017 + HCVData$VISIT_COUNT_2018
# 
#  age <- function(dob, curryear, units = "years", floor = TRUE) {
#      calc.age = interval(dob, curryear) / duration(num = 1, units = units)
#      return(as.integer(floor(calc.age)))
#  }
# 
# HCVData$AGE_2009 <- age(dob = HCVData$BIRTH_DATE, curryear = as.Date('2009-12-31'))
# 
# HCVData$AGE_2010 <- age(dob = HCVData$BIRTH_DATE, curryear = as.Date('2010-12-31'))
# 
# HCVData$AGE_2011 <- age(dob = HCVData$BIRTH_DATE, curryear = as.Date('2011-12-31'))
# 
# HCVData$AGE_2012 <- age(dob = HCVData$BIRTH_DATE, curryear = as.Date('2012-12-31'))
# 
# HCVData$AGE_2013 <- age(dob = HCVData$BIRTH_DATE, curryear = as.Date('2013-12-31'))
# 
# HCVData$AGE_2014 <- age(dob = HCVData$BIRTH_DATE, curryear = as.Date('2014-12-31'))
# 
# HCVData$AGE_2015 <- age(dob = HCVData$BIRTH_DATE, curryear = as.Date('2015-12-31'))
# 
# HCVData$AGE_2016 <- age(dob = HCVData$BIRTH_DATE, curryear = as.Date('2016-12-31'))
# 
# HCVData$AGE_2017 <- age(dob = HCVData$BIRTH_DATE, curryear = as.Date('2017-12-31'))
# 
# HCVData$AGE_2018 <- age(dob = HCVData$BIRTH_DATE, curryear = as.Date('2018-12-31'))
# 
# HCVData$AGE_2019 <- age(dob = HCVData$BIRTH_DATE, curryear = as.Date('2019-12-31'))
# 
# #drop DOB from dataset for privacy reasons
# HCVData <- subset(HCVData, select = -c(BIRTH_DATE))
# 
# HCVData$FIB4_2009 <- HCVData$AGE_2009 * HCVData$AST_2009 / (HCVData$PLATELET_2009 * sqrt(HCVData$ALT_2009))
# 
# HCVData$FIB4_2010 <- HCVData$AGE_2010 * HCVData$AST_2010 / (HCVData$PLATELET_2010 * sqrt(HCVData$ALT_2010))
# 
# HCVData$FIB4_2011 <- HCVData$AGE_2011 * HCVData$AST_2011 / (HCVData$PLATELET_2011 * sqrt(HCVData$ALT_2011))
# 
# HCVData$FIB4_2012 <- HCVData$AGE_2012 * HCVData$AST_2012 / (HCVData$PLATELET_2012 * sqrt(HCVData$ALT_2012))
# 
# HCVData$FIB4_2013 <- HCVData$AGE_2013 * HCVData$AST_2013 / (HCVData$PLATELET_2013 * sqrt(HCVData$ALT_2013))
# 
# HCVData$FIB4_2014 <- HCVData$AGE_2011 * HCVData$AST_2014 / (HCVData$PLATELET_2014 * sqrt(HCVData$ALT_2014))
# 
# HCVData$FIB4_2015 <- HCVData$AGE_2011 * HCVData$AST_2015 / (HCVData$PLATELET_2015 * sqrt(HCVData$ALT_2015))
# 
# HCVData$FIB4_2016 <- HCVData$AGE_2011 * HCVData$AST_2016 / (HCVData$PLATELET_2016 * sqrt(HCVData$ALT_2016))
# 
# HCVData$FIB4_2017 <- HCVData$AGE_2011 * HCVData$AST_2017 / (HCVData$PLATELET_2017 * sqrt(HCVData$ALT_2017))
# 
# HCVData$FIB4_2018 <- HCVData$AGE_2011 * HCVData$AST_2018 / (HCVData$PLATELET_2018 * sqrt(HCVData$ALT_2018))
# 
# HCVData$FIB4_2019 <- HCVData$AGE_2011 * HCVData$AST_2019 / (HCVData$PLATELET_2019 * sqrt(HCVData$ALT_2019))
# 
# save(HCVData, file="HCVData.Rdata")

```

```{r filter_to_adults_with_visits_and_no_Ca_at_start_historical_FIB4s, include = FALSE, eval = TRUE}
load("HCVData.Rdata") 

#consolidating race categories
HCVData$race.ethnicity <- ifelse(HCVData$Race == "White or Caucasian" & HCVData$Ethnicity == "Not Hispanic or Latino", "W", ifelse(HCVData$Race == "Black or African American" & HCVData$Ethnicity == "Not Hispanic or Latino", "B", ifelse(HCVData$Ethnicity == "Hispanic or Latino", "L", ifelse(HCVData$Race == "Unknown" | HCVData$Race == "Declined" | HCVData$Race == "Unknown/Declined", NA, "O"))))
 
# #calculating a total fib4, NOT  because it's meaningful but to be ableto see how many had NO FIB4
# #commenting out for now bc just used this to explore
# HCVData$totalFIB4 <-rowSums(cbind(HCVData$FIB4_2009,
#                                   HCVData$FIB4_2010,
#                                   HCVData$FIB4_2011,
#                                   HCVData$FIB4_2012,
#                                   HCVData$FIB4_2013,
#                                   HCVData$FIB4_2014,
#                                   HCVData$FIB4_2015,
#                                   HCVData$FIB4_2016,
#                                   HCVData$FIB4_2017,
#                                   HCVData$FIB4_2018), na.rm=TRUE)
# summary(HCVData$totalFIB4==0)
# 
# HCVData$totalFIB4BY2011 <-rowSums(cbind(HCVData$FIB4_2009,
#                                   HCVData$FIB4_2010,
#                                   HCVData$FIB4_2011), na.rm=TRUE)
# summary(HCVData$totalFIB4BY2011==0)
# 
# HCVData$totalFIB4BY2014 <-rowSums(cbind(HCVData$FIB4_2009,
#                                   HCVData$FIB4_2010,
#                                   HCVData$FIB4_2011,
#                                   HCVData$FIB4_2012,
#                                   HCVData$FIB4_2013,
#                                   HCVData$FIB4_2014), na.rm=TRUE)
# summary(HCVData$totalFIB4BY2014==0)
# 
# HCVData$totalFIB4BY2015 <-rowSums(cbind(HCVData$FIB4_2009,
#                                   HCVData$FIB4_2010,
#                                   HCVData$FIB4_2011,
#                                   HCVData$FIB4_2012,
#                                   HCVData$FIB4_2013,
#                                   HCVData$FIB4_2014,
#                                   HCVData$FIB4_2015), na.rm=TRUE)
# summary(HCVData$totalFIB4BY2015==0)
# #note- by 2015 72% of our sample has a FIB4.
# 
# HCVData$totalFIB4BY2016 <-rowSums(cbind(HCVData$FIB4_2009,
#                                   HCVData$FIB4_2010,
#                                   HCVData$FIB4_2011,
#                                   HCVData$FIB4_2012,
#                                   HCVData$FIB4_2013,
#                                   HCVData$FIB4_2014,
#                                   HCVData$FIB4_2015,
#                                   HCVData$FIB4_2016), na.rm=TRUE)
# summary(HCVData$totalFIB4BY2016==0)
# 
# HCVData$totalFIB4BY2017 <-rowSums(cbind(HCVData$FIB4_2009,
#                                   HCVData$FIB4_2010,
#                                   HCVData$FIB4_2011,
#                                   HCVData$FIB4_2012,
#                                   HCVData$FIB4_2013,
#                                   HCVData$FIB4_2014,
#                                   HCVData$FIB4_2015,
#                                   HCVData$FIB4_2016,
#                                   HCVData$FIB4_2017), na.rm=TRUE)
# summary(HCVData$totalFIB4BY2017==0)
# 
# HCVData$totalFIB4BY2018 <-rowSums(cbind(HCVData$FIB4_2009,
#                                   HCVData$FIB4_2010,
#                                   HCVData$FIB4_2011,
#                                   HCVData$FIB4_2012,
#                                   HCVData$FIB4_2013,
#                                   HCVData$FIB4_2014,
#                                   HCVData$FIB4_2015,
#                                   HCVData$FIB4_2016,
#                                   HCVData$FIB4_2017,
#                                   HCVData$FIB4_2018), na.rm=TRUE)
# summary(HCVData$totalFIB4BY2018==0)
# 
# HCVData$totalFIB4BY2019 <-rowSums(cbind(HCVData$FIB4_2009,
#                                   HCVData$FIB4_2010,
#                                   HCVData$FIB4_2011,
#                                   HCVData$FIB4_2012,
#                                   HCVData$FIB4_2013,
#                                   HCVData$FIB4_2014,
#                                   HCVData$FIB4_2015,
#                                   HCVData$FIB4_2016,
#                                   HCVData$FIB4_2017,
#                                   HCVData$FIB4_2018,
#                                   HCVData$FIB4_2019), na.rm=TRUE)
# summary(HCVData$totalFIB4BY2019==0)

#Pulling forward FIB4 scores until 2014 where needed
HCVData <- HCVData %>% mutate(FIB4_by2010 = FIB4_2010, YearsSinceFib4_2010 = 0)
HCVData <- HCVData %>% mutate(FIB4_by2010 = ifelse(is.na(FIB4_2010)== "TRUE", FIB4_2009, FIB4_by2010),
                               YearsSinceFib4_2010 = ifelse(is.na(FIB4_2010)== "TRUE", 1, 0))
 
HCVData <-  HCVData %>% mutate(FIB4_by2011 = FIB4_2011, YearsSinceFib4_2011 = 0)
HCVData <-  HCVData %>% mutate(FIB4_by2011 = ifelse(is.na(FIB4_2011)== "TRUE", FIB4_by2010, FIB4_by2011),
                               YearsSinceFib4_2011 = ifelse(is.na(FIB4_2011)== "TRUE", YearsSinceFib4_2010 + 1, 0))

HCVData <-  HCVData %>% mutate(FIB4_by2012 = FIB4_2012, YearsSinceFib4_2012 = 0)
HCVData <-  HCVData %>% mutate(FIB4_by2012 = ifelse(is.na(FIB4_2012)== "TRUE", FIB4_by2011, FIB4_by2012),
                               YearsSinceFib4_2012 = ifelse(is.na(FIB4_2012)== "TRUE", YearsSinceFib4_2011 + 1, 0))

HCVData <-  HCVData %>% mutate(FIB4_by2013 = FIB4_2013, YearsSinceFib4_2013 = 0)
HCVData <-  HCVData %>%  mutate(FIB4_by2013 = ifelse(is.na(FIB4_2013)== "TRUE", FIB4_by2012, FIB4_by2013),
                               YearsSinceFib4_2013 = ifelse(is.na(FIB4_2013)== "TRUE", YearsSinceFib4_2012 + 1, 0))

HCVData <-  HCVData %>% mutate(FIB4_by2014 = FIB4_2014, YearsSinceFib4_2014 = 0)
HCVData <-  HCVData %>% mutate(FIB4_by2014 = ifelse(is.na(FIB4_2014)== "TRUE", FIB4_by2013, FIB4_by2014),
                               YearsSinceFib4_2014 = ifelse(is.na(FIB4_2014)== "TRUE", YearsSinceFib4_2013 + 1, 0))

#Fixing Outcome Variables
HCVData$HCC_2012only <- ifelse((HCVData$Cancer.Diagnosis.Date<'2012-1-1' | HCVData$Cancer.Diagnosis.Date>'2012-12-31') | is.na(HCVData$Cancer.Diagnosis.Date), 0, 1)
HCVData$HCC_2013only <- ifelse((HCVData$Cancer.Diagnosis.Date<'2013-1-1' | HCVData$Cancer.Diagnosis.Date>'2013-12-31') | is.na(HCVData$Cancer.Diagnosis.Date), 0, 1)
HCVData$HCC_2014only <- ifelse((HCVData$Cancer.Diagnosis.Date<'2014-1-1' | HCVData$Cancer.Diagnosis.Date>'2014-12-31') | is.na(HCVData$Cancer.Diagnosis.Date), 0, 1)
HCVData$HCC_2015only <- ifelse((HCVData$Cancer.Diagnosis.Date<'2015-1-1' | HCVData$Cancer.Diagnosis.Date>'2015-12-31') | is.na(HCVData$Cancer.Diagnosis.Date), 0, 1)
HCVData$HCC_2016only <- ifelse((HCVData$Cancer.Diagnosis.Date<'2016-1-1' | HCVData$Cancer.Diagnosis.Date>'2016-12-31') | is.na(HCVData$Cancer.Diagnosis.Date), 0, 1)
HCVData$HCC_2017only <- ifelse((HCVData$Cancer.Diagnosis.Date<'2017-1-1' | HCVData$Cancer.Diagnosis.Date>'2017-12-31') | is.na(HCVData$Cancer.Diagnosis.Date), 0, 1)
HCVData$HCC_2018only <- ifelse((HCVData$Cancer.Diagnosis.Date<'2018-1-1' | HCVData$Cancer.Diagnosis.Date>'2018-12-31') | is.na(HCVData$Cancer.Diagnosis.Date), 0, 1)
HCVData$HCC_2019only <- ifelse((HCVData$Cancer.Diagnosis.Date<'2019-1-1' | HCVData$Cancer.Diagnosis.Date>'2019-12-31') | is.na(HCVData$Cancer.Diagnosis.Date), 0, 1)

HCVData$HCC_2012 <- ifelse(HCVData$Cancer.Diagnosis.Date>'2012-12-31' | is.na(HCVData$Cancer.Diagnosis.Date), 0, 1)
HCVData$HCC_2013 <- ifelse(HCVData$Cancer.Diagnosis.Date>'2013-12-31' | is.na(HCVData$Cancer.Diagnosis.Date), 0, 1)
HCVData$HCC_2014 <- ifelse(HCVData$Cancer.Diagnosis.Date>'2014-12-31' | is.na(HCVData$Cancer.Diagnosis.Date), 0, 1)
HCVData$HCC_2015 <- ifelse(HCVData$Cancer.Diagnosis.Date>'2015-12-31' | is.na(HCVData$Cancer.Diagnosis.Date), 0, 1)
HCVData$HCC_2016 <- ifelse(HCVData$Cancer.Diagnosis.Date>'2016-12-31' | is.na(HCVData$Cancer.Diagnosis.Date), 0, 1)
HCVData$HCC_2017 <- ifelse(HCVData$Cancer.Diagnosis.Date>'2017-12-31' | is.na(HCVData$Cancer.Diagnosis.Date), 0, 1)
HCVData$HCC_2018 <- ifelse(HCVData$Cancer.Diagnosis.Date>'2018-12-31' | is.na(HCVData$Cancer.Diagnosis.Date), 0, 1)
HCVData$HCC_2019 <- ifelse(HCVData$Cancer.Diagnosis.Date>'2019-12-31' | is.na(HCVData$Cancer.Diagnosis.Date), 0, 1)

HCVData$InSystemBy2015 <- ifelse(as.POSIXct(HCVData$First.Hep.Appt) <'2015-12-31' | 
                                   as.POSIXct(HCVData$FIRST_PC_APPT) <'2015-12-31' | 
                                   as.POSIXct(HCVData$First.Hep.Appt) <'2015-12-31' | 
                                   as.POSIXct(HCVData$date.first.pos.HCV.test) < '2015-12-31' | 
                                   as.POSIXct(HCVData$Date.First.Pos.RNA) < '2015-12-31' | 
                                   as.POSIXct(HCVData$AFP.Date) < '2015-12-31'| 
                                   as.POSIXct(HCVData$Lowest.Platelet.Date) < '2015-12-31'| 
                                   as.POSIXct(HCVData$APPT_DATE) < '2015-12-31' | 
                                   as.POSIXct(HCVData$HBsAG.Date) < '2015-12-31' | 
                                   as.POSIXct(HCVData$HBsAg_Date) < '2015-12-31'| 
                                   as.POSIXct(HCVData$HIV.Date) < '2015-12-31'|
                                   as.POSIXct(HCVData$HepA_IgG.Date) < '2015-12-31'|
                                   as.POSIXct(HCVData$HEPB_CORE_IgM_Date) < '2015-12-31' |
                                   as.POSIXct(HCVData$HEPB_CORE_Total_Date) < '2015-12-31' |
                                   as.POSIXct(HCVData$ABD_IMAGING_DATE) < '2015-12-31' |
                                   as.POSIXct(HCVData$BIOPSY_DATE) < '2015-12-31' |
                                   HCVData$VISIT_COUNT_2011 >0 |
                                   HCVData$VISIT_COUNT_2012 >0 |
                                   HCVData$VISIT_COUNT_2013 >0 |
                                   HCVData$VISIT_COUNT_2014 >0 |
                                   HCVData$VISIT_COUNT_2015 >0 |
                                   !is.na(HCVData$AST_2009) |
                                   !is.na(HCVData$AST_2010) |
                                   !is.na(HCVData$AST_2011) |
                                   !is.na(HCVData$AST_2012) |
                                   !is.na(HCVData$AST_2013) |
                                   !is.na(HCVData$AST_2014) |
                                   !is.na(HCVData$AST_2015) |
                                   !is.na(HCVData$ALT_2009) |
                                   !is.na(HCVData$ALT_2010) |
                                   !is.na(HCVData$ALT_2011) |
                                   !is.na(HCVData$ALT_2012) |
                                   !is.na(HCVData$ALT_2013) |
                                   !is.na(HCVData$ALT_2014) |
                                   !is.na(HCVData$ALT_2015) |
                                   !is.na(HCVData$PLATELET_2009) |
                                   !is.na(HCVData$PLATELET_2010) |
                                   !is.na(HCVData$PLATELET_2011) |
                                   !is.na(HCVData$PLATELET_2012) |
                                   !is.na(HCVData$PLATELET_2013) |
                                   !is.na(HCVData$PLATELET_2014) |
                                   !is.na(HCVData$PLATELET_2015), 1, 0)

#Filtering to those who had some visits in 2015-2019, were adults at the start of 2015 and didn't have HCC before 12/31/14
HCVData <- HCVData %>% filter(TotalVisits != 0) %>%
                      filter(AGE_2015 >18) %>%
                      filter(HCC_2014 == 0) %>%
                      filter(InSystemBy2015 ==1)

#Changing abdominal imaging variables to be zero if it's NA
HCVData$ABD_IMAGING_2015[is.na(HCVData$ABD_IMAGING_2015)] <-0
HCVData$ABD_IMAGING_2016[is.na(HCVData$ABD_IMAGING_2016)] <-0
HCVData$ABD_IMAGING_2017[is.na(HCVData$ABD_IMAGING_2017)] <-0
HCVData$ABD_IMAGING_2018[is.na(HCVData$ABD_IMAGING_2018)] <-0

#Making Abd imaging variables that are yes/no to whether someone had imaging that year
HCVData$ABD_IMAGING_YN_2015[HCVData$ABD_IMAGING_2015 == 0] <- 0
HCVData$ABD_IMAGING_YN_2015[HCVData$ABD_IMAGING_2015 > 0] <- 1

HCVData$ABD_IMAGING_YN_2016[HCVData$ABD_IMAGING_2016 == 0] <- 0
HCVData$ABD_IMAGING_YN_2016[HCVData$ABD_IMAGING_2016 > 0] <- 1

HCVData$ABD_IMAGING_YN_2017[HCVData$ABD_IMAGING_2017 == 0] <- 0
HCVData$ABD_IMAGING_YN_2017[HCVData$ABD_IMAGING_2017 > 0] <- 1

HCVData$ABD_IMAGING_YN_2018[HCVData$ABD_IMAGING_2018 == 0] <- 0
HCVData$ABD_IMAGING_YN_2018[HCVData$ABD_IMAGING_2018 > 0] <- 1

#These were just to explore the data once we filtered out those with cirrhosis at baseline
# table(HCVData_inclcirr$HCC_2019)
# table(HCVData$HCC_2019)
# HCVData_outcome <- HCVData %>% filter(HCC_2019==1)
# HCVData_nooutcome <- HCVData %>% filter(HCC_2019==0)
# summary(HCVData_outcome$FIB4_2019)
# summary(HCVData_nooutcome$FIB4_2019)
# 
# table(HCVData$VISIT_COUNT_2015, HCVData$ABD_IMAGING_2015)
# 
# HCVData_imaging <- HCVData %>% filter(HCVData$ABD_IMAGING_2015 > 0 & HCVData$ABD_IMAGING_2016 > 0 & HCVData$ABD_IMAGING_2017 > 0 & HCVData$ABD_IMAGING_2018 > 0)
# table(HCVData_imaging$HCC_2019)


#impute FIB4, SES, Race
HCVData$PayorType <- as.factor(HCVData$PayorType)
HCVData$race.ethnicity <- as.factor(HCVData$race.ethnicity)
impute <- c("PtID", "PayorType", "race.ethnicity", "FIB4_by2014")
keeps <- c("PayorType", "race.ethnicity", "FIB4_by2014")

###commenting out imputation to save computation time###
#imputation <- mice(HCVData[, (names(HCVData) %in% impute)], m=5, maxit=5, nnet.MaxNWts=10000, printFlag=TRUE, seed = 252)
load("ImputedData.Rdata")

impute1 <- complete(imputation, 1)
impute2 <- complete(imputation, 2)
impute3 <- complete(imputation, 3)
impute4 <- complete(imputation, 4)
impute5 <- complete(imputation, 5)
HCVData1 <- merge(impute1, HCVData[, !(names(HCVData) %in% keeps)], by = "PtID")
HCVData2 <- merge(impute2, HCVData[, !(names(HCVData) %in% keeps)], by = "PtID")
HCVData3 <- merge(impute3, HCVData[, !(names(HCVData) %in% keeps)], by = "PtID")
HCVData4 <- merge(impute4, HCVData[, !(names(HCVData) %in% keeps)], by = "PtID")
HCVData5 <- merge(impute5, HCVData[, !(names(HCVData) %in% keeps)], by = "PtID")

# imputed <- rbindlist(list(impute1, impute2, impute3, impute4, impute5))[,lapply(.SD,mean), list(PtID)]
# 
# calculate_mode <- function(x) {
#   uniqx <- unique(na.omit(x))
#   uniqx[which.max(tabulate(match(x, uniqx)))]
# }
# race.ethnicity_imp <- mode()

# #OLD - keeping just in case
# #Imputing as the median value for those without a FIB4 by 2014
# HCVData <-  HCVData %>% mutate(YearsSinceFib4_2014 = ifelse(is.na(FIB4_by2014)== "TRUE", 0, YearsSinceFib4_2014),
#                                ImputedFIB4_2014 = ifelse(is.na(FIB4_by2014)== "TRUE", 1, 0),
#                                FIB4_by2014 = ifelse(is.na(FIB4_by2014)== "TRUE", median(HCVData$FIB4_by2014,na.rm= "TRUE"), FIB4_by2014))
# 
# HCVData <-  HCVData %>% mutate(FIB4_by2015 = FIB4_2015, YearsSinceFib4_2015 = 0)
# HCVData <-  HCVData %>% mutate(FIB4_by2015 = ifelse(is.na(FIB4_2015)== "TRUE", FIB4_by2014, FIB4_by2015),
#                                YearsSinceFib4_2015 = ifelse(is.na(FIB4_2015)== "TRUE", YearsSinceFib4_2014 + 1, 0),
#                                ImputedFIB4_2015 = ifelse(is.na(FIB4_2015)== "TRUE", ImputedFIB4_2014, 0))
# 
# HCVData <-  HCVData %>% mutate(FIB4_by2016 = FIB4_2016, YearsSinceFib4_2016 = 0)
# HCVData <-  HCVData %>% mutate(FIB4_by2016 = ifelse(is.na(FIB4_2016)== "TRUE", FIB4_by2015, FIB4_by2016),
#                                YearsSinceFib4_2016 = ifelse(is.na(FIB4_2016)== "TRUE", YearsSinceFib4_2015 + 1, 0),
#                                ImputedFIB4_2016 = ifelse(is.na(FIB4_2016)== "TRUE", ImputedFIB4_2015, 0))
# 
# HCVData <-  HCVData %>% mutate(FIB4_by2017 = FIB4_2017, YearsSinceFib4_2017 = 0)
# HCVData <-  HCVData %>% mutate(FIB4_by2017 = ifelse(is.na(FIB4_2017)== "TRUE", FIB4_by2016, FIB4_by2017),
#                                YearsSinceFib4_2017 = ifelse(is.na(FIB4_2017)== "TRUE", YearsSinceFib4_2016 + 1, 0),
#                                ImputedFIB4_2017 = ifelse(is.na(FIB4_2017)== "TRUE", ImputedFIB4_2016, 0))
# 
# HCVData <-  HCVData %>% mutate(FIB4_by2018 = FIB4_2018, YearsSinceFib4_2018 = 0)
# HCVData <-  HCVData %>% mutate(FIB4_by2018 = ifelse(is.na(FIB4_2018)== "TRUE", FIB4_by2017, FIB4_by2018),
#                                YearsSinceFib4_2018 = ifelse(is.na(FIB4_2018)== "TRUE", YearsSinceFib4_2017 + 1, 0),
#                                ImputedFIB4_2018 = ifelse(is.na(FIB4_2018)== "TRUE", ImputedFIB4_2017, 0))
# 
# HCVData <-  HCVData %>% mutate(FIB4_by2019 = FIB4_2019, YearsSinceFib4_2019 = 0)
# HCVData <-  HCVData %>% mutate(FIB4_by2019 = ifelse(is.na(FIB4_2019)== "TRUE", FIB4_by2018, FIB4_by2019),
#                                YearsSinceFib4_2019 = ifelse(is.na(FIB4_2019)== "TRUE", YearsSinceFib4_2018 + 1, 0),
#                                ImputedFIB4_2019 = ifelse(is.na(FIB4_2019)== "TRUE", ImputedFIB4_2018, 0))

#Filtering to those with no cirrhosis at baseline
HCVData1 <- HCVData1 %>% filter(FIB4_by2014 < 3.25)
HCVData2 <- HCVData2 %>% filter(FIB4_by2014 < 3.25)
HCVData3 <- HCVData3 %>% filter(FIB4_by2014 < 3.25)
HCVData4 <- HCVData4 %>% filter(FIB4_by2014 < 3.25)
HCVData5 <- HCVData5 %>% filter(FIB4_by2014 < 3.25)

#Pulling forward FIB4s post-2014
HCVData1 <-  HCVData1 %>% mutate(FIB4_by2015 = FIB4_2015, YearsSinceFib4_2015 = 0)
HCVData1 <-  HCVData1 %>% mutate(
  FIB4_by2015 = ifelse(is.na(FIB4_2015)== "TRUE", FIB4_by2014, FIB4_by2015),
  YearsSinceFib4_2015 = ifelse(is.na(FIB4_2015)== "TRUE", YearsSinceFib4_2014 + 1, 0),
  ImputedFIB4_2015 = ifelse(is.na(FIB4_2015)== "TRUE", FIB4_by2014, 0))
HCVData1 <-  HCVData1 %>% mutate(FIB4_by2016 = FIB4_2016, YearsSinceFib4_2016 = 0)
HCVData1 <-  HCVData1 %>% mutate(FIB4_by2016 = ifelse(is.na(FIB4_2016)== "TRUE", FIB4_by2015, FIB4_by2016),
                               YearsSinceFib4_2016 = ifelse(is.na(FIB4_2016)== "TRUE", YearsSinceFib4_2015 + 1, 0),
                               ImputedFIB4_2016 = ifelse(is.na(FIB4_2016)== "TRUE", ImputedFIB4_2015, 0))
HCVData1 <-  HCVData1 %>% mutate(FIB4_by2017 = FIB4_2017, YearsSinceFib4_2017 = 0)
HCVData1 <-  HCVData1 %>% mutate(FIB4_by2017 = ifelse(is.na(FIB4_2017)== "TRUE", FIB4_by2016, FIB4_by2017),
                               YearsSinceFib4_2017 = ifelse(is.na(FIB4_2017)== "TRUE", YearsSinceFib4_2016 + 1, 0),
                               ImputedFIB4_2017 = ifelse(is.na(FIB4_2017)== "TRUE", ImputedFIB4_2016, 0))
HCVData1 <-  HCVData1 %>% mutate(FIB4_by2018 = FIB4_2018, YearsSinceFib4_2018 = 0)
HCVData1 <-  HCVData1 %>% mutate(FIB4_by2018 = ifelse(is.na(FIB4_2018)== "TRUE", FIB4_by2017, FIB4_by2018),
                               YearsSinceFib4_2018 = ifelse(is.na(FIB4_2018)== "TRUE", YearsSinceFib4_2017 + 1, 0),
                               ImputedFIB4_2018 = ifelse(is.na(FIB4_2018)== "TRUE", ImputedFIB4_2017, 0))
HCVData1 <-  HCVData1 %>% mutate(FIB4_by2019 = FIB4_2019, YearsSinceFib4_2019 = 0)
HCVData1 <-  HCVData1 %>% mutate(FIB4_by2019 = ifelse(is.na(FIB4_2019)== "TRUE", FIB4_by2018, FIB4_by2019),
                               YearsSinceFib4_2019 = ifelse(is.na(FIB4_2019)== "TRUE", YearsSinceFib4_2018 + 1, 0),
                               ImputedFIB4_2019 = ifelse(is.na(FIB4_2019)== "TRUE", ImputedFIB4_2018, 0))

HCVData2 <-  HCVData2 %>% mutate(FIB4_by2015 = FIB4_2015, YearsSinceFib4_2015 = 0)
HCVData2 <-  HCVData2 %>% mutate(FIB4_by2015 = ifelse(is.na(FIB4_2015)== "TRUE", FIB4_by2014, FIB4_by2015),
                               YearsSinceFib4_2015 = ifelse(is.na(FIB4_2015)== "TRUE", YearsSinceFib4_2014 + 1, 0),
                               ImputedFIB4_2015 = ifelse(is.na(FIB4_2015)== "TRUE", FIB4_by2014, 0))
HCVData2 <-  HCVData2 %>% mutate(FIB4_by2016 = FIB4_2016, YearsSinceFib4_2016 = 0)
HCVData2 <-  HCVData2 %>% mutate(FIB4_by2016 = ifelse(is.na(FIB4_2016)== "TRUE", FIB4_by2015, FIB4_by2016),
                               YearsSinceFib4_2016 = ifelse(is.na(FIB4_2016)== "TRUE", YearsSinceFib4_2015 + 1, 0),
                               ImputedFIB4_2016 = ifelse(is.na(FIB4_2016)== "TRUE", ImputedFIB4_2015, 0))
HCVData2 <-  HCVData2 %>% mutate(FIB4_by2017 = FIB4_2017, YearsSinceFib4_2017 = 0)
HCVData2 <-  HCVData2 %>% mutate(FIB4_by2017 = ifelse(is.na(FIB4_2017)== "TRUE", FIB4_by2016, FIB4_by2017),
                               YearsSinceFib4_2017 = ifelse(is.na(FIB4_2017)== "TRUE", YearsSinceFib4_2016 + 1, 0),
                               ImputedFIB4_2017 = ifelse(is.na(FIB4_2017)== "TRUE", ImputedFIB4_2016, 0))
HCVData2 <-  HCVData2 %>% mutate(FIB4_by2018 = FIB4_2018, YearsSinceFib4_2018 = 0)
HCVData2 <-  HCVData2 %>% mutate(FIB4_by2018 = ifelse(is.na(FIB4_2018)== "TRUE", FIB4_by2017, FIB4_by2018),
                               YearsSinceFib4_2018 = ifelse(is.na(FIB4_2018)== "TRUE", YearsSinceFib4_2017 + 1, 0),
                               ImputedFIB4_2018 = ifelse(is.na(FIB4_2018)== "TRUE", ImputedFIB4_2017, 0))
HCVData2 <-  HCVData2 %>% mutate(FIB4_by2019 = FIB4_2019, YearsSinceFib4_2019 = 0)
HCVData2 <-  HCVData2 %>% mutate(FIB4_by2019 = ifelse(is.na(FIB4_2019)== "TRUE", FIB4_by2018, FIB4_by2019),
                               YearsSinceFib4_2019 = ifelse(is.na(FIB4_2019)== "TRUE", YearsSinceFib4_2018 + 1, 0),
                               ImputedFIB4_2019 = ifelse(is.na(FIB4_2019)== "TRUE", ImputedFIB4_2018, 0))

HCVData3 <-  HCVData3 %>% mutate(FIB4_by2015 = FIB4_2015, YearsSinceFib4_2015 = 0)
HCVData3 <-  HCVData3 %>% mutate(FIB4_by2015 = ifelse(is.na(FIB4_2015)== "TRUE", FIB4_by2014, FIB4_by2015),
                               YearsSinceFib4_2015 = ifelse(is.na(FIB4_2015)== "TRUE", YearsSinceFib4_2014 + 1, 0),
                               ImputedFIB4_2015 = ifelse(is.na(FIB4_2015)== "TRUE", FIB4_by2014, 0))
HCVData3 <-  HCVData3 %>% mutate(FIB4_by2016 = FIB4_2016, YearsSinceFib4_2016 = 0)
HCVData3 <-  HCVData3 %>% mutate(FIB4_by2016 = ifelse(is.na(FIB4_2016)== "TRUE", FIB4_by2015, FIB4_by2016),
                               YearsSinceFib4_2016 = ifelse(is.na(FIB4_2016)== "TRUE", YearsSinceFib4_2015 + 1, 0),
                               ImputedFIB4_2016 = ifelse(is.na(FIB4_2016)== "TRUE", ImputedFIB4_2015, 0))
HCVData3 <-  HCVData3 %>% mutate(FIB4_by2017 = FIB4_2017, YearsSinceFib4_2017 = 0)
HCVData3 <-  HCVData3 %>% mutate(FIB4_by2017 = ifelse(is.na(FIB4_2017)== "TRUE", FIB4_by2016, FIB4_by2017),
                               YearsSinceFib4_2017 = ifelse(is.na(FIB4_2017)== "TRUE", YearsSinceFib4_2016 + 1, 0),
                               ImputedFIB4_2017 = ifelse(is.na(FIB4_2017)== "TRUE", ImputedFIB4_2016, 0))
HCVData3 <-  HCVData3 %>% mutate(FIB4_by2018 = FIB4_2018, YearsSinceFib4_2018 = 0)
HCVData3 <-  HCVData3 %>% mutate(FIB4_by2018 = ifelse(is.na(FIB4_2018)== "TRUE", FIB4_by2017, FIB4_by2018),
                               YearsSinceFib4_2018 = ifelse(is.na(FIB4_2018)== "TRUE", YearsSinceFib4_2017 + 1, 0),
                               ImputedFIB4_2018 = ifelse(is.na(FIB4_2018)== "TRUE", ImputedFIB4_2017, 0))
HCVData3 <-  HCVData3 %>% mutate(FIB4_by2019 = FIB4_2019, YearsSinceFib4_2019 = 0)
HCVData3 <-  HCVData3 %>% mutate(FIB4_by2019 = ifelse(is.na(FIB4_2019)== "TRUE", FIB4_by2018, FIB4_by2019),
                               YearsSinceFib4_2019 = ifelse(is.na(FIB4_2019)== "TRUE", YearsSinceFib4_2018 + 1, 0),
                               ImputedFIB4_2019 = ifelse(is.na(FIB4_2019)== "TRUE", ImputedFIB4_2018, 0))

HCVData4 <-  HCVData4 %>% mutate(FIB4_by2015 = FIB4_2015, YearsSinceFib4_2015 = 0)
HCVData4 <-  HCVData4 %>% mutate(FIB4_by2015 = ifelse(is.na(FIB4_2015)== "TRUE", FIB4_by2014, FIB4_by2015),
                               YearsSinceFib4_2015 = ifelse(is.na(FIB4_2015)== "TRUE", YearsSinceFib4_2014 + 1, 0),
                               ImputedFIB4_2015 = ifelse(is.na(FIB4_2015)== "TRUE", FIB4_by2014, 0))
HCVData4 <-  HCVData4 %>% mutate(FIB4_by2016 = FIB4_2016, YearsSinceFib4_2016 = 0)
HCVData4 <-  HCVData4 %>% mutate(FIB4_by2016 = ifelse(is.na(FIB4_2016)== "TRUE", FIB4_by2015, FIB4_by2016),
                               YearsSinceFib4_2016 = ifelse(is.na(FIB4_2016)== "TRUE", YearsSinceFib4_2015 + 1, 0),
                               ImputedFIB4_2016 = ifelse(is.na(FIB4_2016)== "TRUE", ImputedFIB4_2015, 0))
HCVData4 <-  HCVData4 %>% mutate(FIB4_by2017 = FIB4_2017, YearsSinceFib4_2017 = 0)
HCVData4 <-  HCVData4 %>% mutate(FIB4_by2017 = ifelse(is.na(FIB4_2017)== "TRUE", FIB4_by2016, FIB4_by2017),
                               YearsSinceFib4_2017 = ifelse(is.na(FIB4_2017)== "TRUE", YearsSinceFib4_2016 + 1, 0),
                               ImputedFIB4_2017 = ifelse(is.na(FIB4_2017)== "TRUE", ImputedFIB4_2016, 0))
HCVData4 <-  HCVData4 %>% mutate(FIB4_by2018 = FIB4_2018, YearsSinceFib4_2018 = 0)
HCVData4 <-  HCVData4 %>% mutate(FIB4_by2018 = ifelse(is.na(FIB4_2018)== "TRUE", FIB4_by2017, FIB4_by2018),
                               YearsSinceFib4_2018 = ifelse(is.na(FIB4_2018)== "TRUE", YearsSinceFib4_2017 + 1, 0),
                               ImputedFIB4_2018 = ifelse(is.na(FIB4_2018)== "TRUE", ImputedFIB4_2017, 0))
HCVData4 <-  HCVData4 %>% mutate(FIB4_by2019 = FIB4_2019, YearsSinceFib4_2019 = 0)
HCVData4 <-  HCVData4 %>% mutate(FIB4_by2019 = ifelse(is.na(FIB4_2019)== "TRUE", FIB4_by2018, FIB4_by2019),
                               YearsSinceFib4_2019 = ifelse(is.na(FIB4_2019)== "TRUE", YearsSinceFib4_2018 + 1, 0),
                               ImputedFIB4_2019 = ifelse(is.na(FIB4_2019)== "TRUE", ImputedFIB4_2018, 0))

HCVData5 <-  HCVData5 %>% mutate(FIB4_by2015 = FIB4_2015, YearsSinceFib4_2015 = 0)
HCVData5 <-  HCVData5 %>% mutate(FIB4_by2015 = ifelse(is.na(FIB4_2015)== "TRUE", FIB4_by2014, FIB4_by2015),
                               YearsSinceFib4_2015 = ifelse(is.na(FIB4_2015)== "TRUE", YearsSinceFib4_2014 + 1, 0),
                               ImputedFIB4_2015 = ifelse(is.na(FIB4_2015)== "TRUE", FIB4_by2014, 0))
HCVData5 <-  HCVData5 %>% mutate(FIB4_by2016 = FIB4_2016, YearsSinceFib4_2016 = 0)
HCVData5 <-  HCVData5 %>% mutate(FIB4_by2016 = ifelse(is.na(FIB4_2016)== "TRUE", FIB4_by2015, FIB4_by2016),
                               YearsSinceFib4_2016 = ifelse(is.na(FIB4_2016)== "TRUE", YearsSinceFib4_2015 + 1, 0),
                               ImputedFIB4_2016 = ifelse(is.na(FIB4_2016)== "TRUE", ImputedFIB4_2015, 0))
HCVData5 <-  HCVData5 %>% mutate(FIB4_by2017 = FIB4_2017, YearsSinceFib4_2017 = 0)
HCVData5 <-  HCVData5 %>% mutate(FIB4_by2017 = ifelse(is.na(FIB4_2017)== "TRUE", FIB4_by2016, FIB4_by2017),
                               YearsSinceFib4_2017 = ifelse(is.na(FIB4_2017)== "TRUE", YearsSinceFib4_2016 + 1, 0),
                               ImputedFIB4_2017 = ifelse(is.na(FIB4_2017)== "TRUE", ImputedFIB4_2016, 0))
HCVData5 <-  HCVData5 %>% mutate(FIB4_by2018 = FIB4_2018, YearsSinceFib4_2018 = 0)
HCVData5 <-  HCVData5 %>% mutate(FIB4_by2018 = ifelse(is.na(FIB4_2018)== "TRUE", FIB4_by2017, FIB4_by2018),
                               YearsSinceFib4_2018 = ifelse(is.na(FIB4_2018)== "TRUE", YearsSinceFib4_2017 + 1, 0),
                               ImputedFIB4_2018 = ifelse(is.na(FIB4_2018)== "TRUE", ImputedFIB4_2017, 0))
HCVData5 <-  HCVData5 %>% mutate(FIB4_by2019 = FIB4_2019, YearsSinceFib4_2019 = 0)
HCVData5 <-  HCVData5 %>% mutate(FIB4_by2019 = ifelse(is.na(FIB4_2019)== "TRUE", FIB4_by2018, FIB4_by2019),
                               YearsSinceFib4_2019 = ifelse(is.na(FIB4_2019)== "TRUE", YearsSinceFib4_2018 + 1, 0),
                               ImputedFIB4_2019 = ifelse(is.na(FIB4_2019)== "TRUE", ImputedFIB4_2018, 0))

#check for years with visits but no imaging
HCVData1$noimage2014 <- ifelse(HCVData1$VISIT_COUNT_2014 > 0 & (HCVData1$ABD_IMAGING_2014==0 | is.na(HCVData1$ABD_IMAGING_2014)), 1, 0)
HCVData1$noimage2015 <- ifelse(HCVData1$VISIT_COUNT_2015 > 0 & (HCVData1$ABD_IMAGING_2015==0 | is.na(HCVData1$ABD_IMAGING_2015)), 1, 0)
HCVData1$noimage2016 <- ifelse(HCVData1$VISIT_COUNT_2016 > 0 & (HCVData1$ABD_IMAGING_2016==0 | is.na(HCVData1$ABD_IMAGING_2016)), 1, 0)
HCVData1$noimage2017 <- ifelse(HCVData1$VISIT_COUNT_2017 > 0 & (HCVData1$ABD_IMAGING_2017==0 | is.na(HCVData1$ABD_IMAGING_2017)), 1, 0)
HCVData1$noimage2018 <- ifelse(HCVData1$VISIT_COUNT_2018 > 0 & (HCVData1$ABD_IMAGING_2018==0 | is.na(HCVData1$ABD_IMAGING_2018)), 1, 0)
HCVData1$noimages <- ifelse((HCVData1$noimage2014 + HCVData1$noimage2015 + HCVData1$noimage2016 + HCVData1$noimage2017 + HCVData1$noimage2018) < 5, 1, 0)
PctNoImages <- round(sum(HCVData1$noimages)/nrow(HCVData1)*100, 1)
                       
```

# 1. Background
## A. Brief Introduction
Hepatocellular carcinoma (HCC), a type of liver cancer, is the second most common cause of liver-related death throughout the world (Bosetti et al 2014). HCC is a major complication of infection with hepatitis C (HCV) virus, occuring in 1%-4% of people with liver cirrhosis each year (Omland et al 2010). HCV becomes chronic infection for about 80% of adults, and if left untreated can cause increasing cirrhosis over a period of 20-30 years of often asymptomatic disease, until damage is severe and major complications are irreversable (El-Serag 2012). Certain HCV viral genotypes (particularly genotype 3) are associated with higher risk of HCC, as is continued smoking (Chuang et al 2010) and alcohol use (Donato et al 2002), as well as concurrent diabetes or obesity (Huang et al 2017, Calle et al 2003). 

Since 2014, direct-acting antiviral (DAA) therapy has dramatically improved prognosis for people living with HCV; 8-12 weeks of well-tolerated oral therapy leads to cure in more than 90% of patients (Burstow et al 2017), halting cirrhosis progression and apparently reducing the elevated risk of HCC (Axley et al 2017). Patients whose HCV has already caused some cirrhosis continue to be at some increased risk of HCC even post-cure, and given the high mortality risk it is recommended that people with HCV infection receive systematic monitoring with liver ultrasound; the European Association for Study of the Liver currently recommends screening twice per year (EASL 2015). However, many UCSF patients chronically infected with HCV do not have evidence of abdominal imaging at least once per year. In this study we aimed to estimate the longitudinal association of abdominal imaging on HCC risk using causal inference methods. We hypothesized that patients with no cirrhosis at baseline would be less likely to develop HCC within 5 years when they had evidence of abdominal imaging at least once per year for 5 years, compared to patients who did not have annual abdominal imaging.

## B. Description of our Dataset

We are using a dataset of chronic hepatitis C patients receiving care in the UCSF system since 2009. There are `r nrow(HCVData)` patients, who were adults seen in the UCSF system by the end of 2015 and have been seen in a variety of primary care clinics and the hepatology (liver) clinic between 2015 and 2019. These data come from a query of Apex, the UCSF-specific build of the electronic medical record system Epic.

We restricted our dataset to adult UCSF patients diagnosed with HCV who, prior to the start of follow up (ie. as of 12/31/14) were not known to have indications of cirrhosis (FIB-4 score > 3.25) or HCC. (n = `r nrow(HCVData1)`) For those who had been in the system prior to 2015, historical data back to 2011 was used to calculate their FIB-4 score at the beinning of follow up (ie. by the end of 2014). For those who did not yet have those data, the FIB-4 score in 2014 was imputed using multiple imputation (m=5). 

<!-- reminder that if someone had no data through 2014, and first showed up in 2015, they're still in this dataset even if they had, say, cirrhosis diagnosed elsewhere prior to 2015. So we should be careful to describe the dataset as those in the system, 'not known to have cirrhosis' prior to 1/1/2015 or sth similar-->

```{r demographics, include = FALSE, eval = TRUE}
#note that the demographics in this table are based on imputation dataset 1, for simplicity
sex <- table(HCVData1$Sex, useNA = "ifany")
race <- table(HCVData1$race.ethnicity, useNA  = "ifany")
SES <- table(HCVData1$PayorType, useNA = "ifany")
n <- nrow(HCVData1)
male <- sex[2]
malep <- male/n*100
female <- sex[1]
femalep <- female/n*100
white <- race[4]
whitep <- white/n*100
black <- race[1]
blackp <- black/n*100
latinx <- race[2]
latinxp <- latinx/n*100
other <- race[3]
otherp <- other/n*100
unkrace <- race[5]
unkracep <- unkrace/n*100
medical <- SES[1]
medicalp <- medical/n*100
notmedical <- SES[2]
notmedicalp <- notmedical/n*100

```

## C. Our Observed Data

**Number of Visits (annually)**

The number of visits per patient ranges from 0 to 44 in any given year, with a median of 1 visit per year overall. The figure below presents number of visits per patient in each year, truncated at 25 visits.

```{r HCC, include=FALSE}
HCVData1$HCC <- ifelse(is.na(HCVData1$Cancer.Diagnosis.Name), 0, 1)
HCCpercent <- mean(HCVData1$HCC)*100
```

**Presence of Abdominal Imaging**

In years where they had at least one visit with a primary care physician or hepatologist, `r PctNoImages`% of UCSF patients chronically infected with HCV had no evidence of abdominal imaging at least one of those years for the 5 years under study.

**Diagnosis of HCC by 2019**

`r sum(HCVData1$HCC)` people (`r round(HCCpercent, 1)`%) developed the outcome of interest (were diagnosed with HCC during the course of the study). The figure below shows the number of people diagnosed with HCC in each year, and cumulatively, throughout the study period (\textbf{note} that 2019 is an incomplete year of data.)


```{r HCCplots, include = TRUE, echo = FALSE, warning = FALSE, message = FALSE, fig.width=11, fig.height=7}

HCCyearly <- HCVData1 %>% dplyr::select(HCC_2016only, HCC_2017only, HCC_2018only, HCC_2019only)
colnames(HCCyearly) <- c("2016", "2017", "2018", "2019")
HCCYearlyLong <- HCCyearly %>% gather
plotHCCann <- ggplot(HCCYearlyLong, aes(x=key, y=value)) + geom_col(fill = "darkgreen") + scale_y_continuous(breaks = seq(0,100,10)) + xlab("Year") + ylab("Number of Patients") + ggtitle("New HCC diagnoses, by year") + theme(legend.position = "none") + theme_minimal() + theme(plot.title = element_text(size = 14, hjust = 0.5))

HCCtotal <- HCVData1 %>% dplyr::select(HCC_2016, HCC_2017, HCC_2018, HCC_2019)
colnames(HCCtotal) <- c("2016", "2017", "2018", "2019")
HCCLong <- HCCtotal %>% gather
plotHCCcum <- ggplot(HCCLong, aes(x=key, y=value)) + geom_col(fill = "darkgreen") + scale_y_continuous(breaks = seq(0,500,50)) + xlab("Year") + ylab("Number of Patients") + ggtitle("Cumulative HCC cases, by year") + theme(legend.position = "none") + theme_minimal() + theme(plot.title = element_text(size = 14, hjust = 0.5))

HCCplots <- ggarrange(plotHCCann, plotHCCcum, ncol = 2, nrow = 1)
suppressWarnings(print(HCCplots))
```

**FIB-4 Score (annually)**

```{r FIBLongitems, include = FALSE, eval = TRUE}
FIBscores <- HCVData1 %>% dplyr::select(FIB4_2011, FIB4_2012, FIB4_2013, FIB4_2014, FIB4_2015, FIB4_2016, FIB4_2017, FIB4_2018)
FIBLong <- FIBscores %>% gather
cirrhosis <- nrow(subset(FIBLong, value>3.25))
outliers <- nrow(subset(FIBLong, value>9))
minFib <- round(range(FIBLong$value, na.rm = TRUE)[1],3)
maxFib <- round(range(FIBLong$value, na.rm = TRUE)[2],2)
FibIQR <- round(IQR(FIBLong$value, na.rm = TRUE),4)
Fib25 <- round(quantile(FIBLong$value, na.rm = TRUE)[2],4)
Fib75 <- round(quantile(FIBLong$value, na.rm = TRUE)[4],4)
```

FIB-4 scores (Sterling, 2006) are calculated with inputs of age, platelet count, AST, and ALT. Scores of <1.45 are considered to be strongly suggestive of no liver fibrosis, and scores >3.25 are indicative of advanced fibrosis and/or cirrhosis. 

FIB-4 scores in this dataset range of `r minFib` to `r maxFib`, with an IQR of `r FibIQR` (`r Fib25`-`r Fib75`), indicating the FIB-4 score of 45 is a substantial outlier. `r cirrhosis` FIB-4 scores are >3.25 throughout all years of follow-up on all patients, indicating likely cirrhosis and increased risk of HCC at those timepoints.  There are `r outliers` instances where a patient's FIB-4 score was calculated to be greater than 9.

Given that FIB-4 scores could not be measured every year for every patient, after 2014 when test results were not available for a new FIB-4 calculation, the prior year's FIB-4 score was carried forward; the number of years since FIB-4 was measured was also included as a longitudinal covariate in our model.
  
\vspace{12pt}

**Other demographics**

\vspace{-6pt}

* \textbf{Sex}. This is a dichotomous variable, with two categories (male, female), as UCSF has not been capturing other gender categories within Epic.
* \textbf{Race}. This is a categorical variable, which recategorized into White, Black/African American, Latinx, and Other for ease of analysis given the bias introduced into causal analyses by practical positivity violations.
* \textbf{SES}. We used insurance type (Medi-Cal, or not Medi-Cal) as a marker of SES status, which is again a dichotomous variable.

The table below displays the demographic breakdown of the sample.

| Demographic | Category | n | % |
| ----------------- | ----------------- | -------- | --------|
| Sex | Male | `r male` | `r round(malep, 1)`% |
|        | Female | `r female` | `r round(femalep, 1)`% |
|  | 
|        | Black/African American | `r black` | `r round(blackp, 1)`% |
|        | Latinx | `r latinx` | `r round(latinxp, 1)`% |
| Race/ethnicity | White | `r white` | `r round(whitep, 1)`% |
|        | Other | `r other` | `r round(otherp, 1)`% |
|        | Unknown | `r unkrace` | `r round(unkracep, 1)`% |  
|  | 
|   SES (Payor type)   | Medi-Cal | `r medical` | `r round(medicalp, 1)`% |
|  | Not Medi-Cal | `r notmedical` | `r round(notmedicalp, 1)`% |
|  | 
| TOTAL | | `r n` | 100% |
 
## D. Structural Model

We defined our **exposure** as the total number of abdominal imaging results reported within the UCSF system for each chronic HCV patient, annually from 2015 through 2018. We then dichotomized this variable (0 images, or 1+ images) due to the current computational limitations of longitudinal targeted machine learning estimation (LTMLE). 

We defined our **outcome** as diagnosis of HCC, by the final year (2019). 

We included 6 **covariates** in our model: race, sex, and SES as defined above, as well as 3 longitudinal covariates assessed annually:

\vspace{-8pt}

* Total number of primary care or hepatology visits in the UCSF system, annually from 2015 - 2019,
* FIB-4 score (a validated measure used for prediction of cirrhosis (Khan et al 2017), which uses age, platelet count and liver transaminases), and
* Years since FIB-4 score was last measured.

This is a survival analysis, with a Structural Causal Model $O= \{W, L(t), Y(t), A(t)\}$ where:

\vspace{-8pt}

* W is the baseline covariates (race, sex and SES)
* L(t) is the set of covariates (number of primary care or hepatology visits, most recent FIB-4 score, years since last FIB-4) at time t
* A(t) is the exposure (presence of abdominal imaging)
* Y(t) is the outcome (an indicator of HCC diagnosis)

\vspace{-2pt}

$U = (U_{L(t)}, U_{A(t)}, U_{C(t)}, U_{Y(t)},), t = 1,2,3,4,5$ ~ $P_U$

\vspace{4pt}

Structural Equations, $\mathcal{M^F}$, for $t$ from 1 to 5, were:
\begin{align*}
W &= f_{W}(U_{W}) \\
L(1) &= f_{L(1)}(U_{L(1)}) \\
A(1) &= f_{A(1)}(W, L(1), U_{A(1)}) \\
L(t) &= f_{L(t)}(\bar{L}(t-1), \bar{A}(t-1), U_{L(t)}) \\
A(t) &= f_{A(t)}(W, \bar{L}(t), \bar{A}(t-1), U_{A(t)}) \\
Y(t) &= f_{Y(t)}(W, \bar{L}(t-1), \bar{A}(t-1), Y(t-1), U_{Y(t)}) \\
Y(K) &= f_{Y(K)}(W, \bar{L}(K-1), \bar{A}(K-1), Y(K-1), U_{Y(K)}) \\
\end{align*}

\vspace{-10pt}

We included one key exclusion restriction: $Y(t)$ is an indicator variable describing whether or not the patient has been diagnosed with HCC by time $t$; as such, once $Y(t)=1$ it is deterministically set to 1 through $Y(K)$. 

```{r Simulation Prep}

#First just exploring our data distributions so our sims can match
#Because A is the number imaging occurrences in a year, A1 is imaging in 2015, A2 is in 2016...
summary(HCVData$ABD_IMAGING_YN_2015)
summary(HCVData$ABD_IMAGING_YN_2016)
summary(HCVData$ABD_IMAGING_YN_2017)
summary(HCVData$ABD_IMAGING_YN_2018)

#Because L_a is visit count, L_a1 is visits in 2015,...
summary(HCVData$VISIT_COUNT_2015)
summary(HCVData$VISIT_COUNT_2016)
summary(HCVData$VISIT_COUNT_2017)
summary(HCVData$VISIT_COUNT_2018)

#Because L_b is fib4 score, L_b1 is FIB4 in 2015,...
summary(HCVData1$FIB4_by2015)
summary(HCVData1$FIB4_by2016)
summary(HCVData1$FIB4_by2017)
summary(HCVData1$FIB4_by2018)

    #In case we're curious, this shows what percent of the FIB4 scores are using the imputed values
    summary(HCVData1$ImputedFIB4_2015)
    summary(HCVData1$ImputedFIB4_2016)
    summary(HCVData1$ImputedFIB4_2017)
    summary(HCVData1$ImputedFIB4_2018)
    #and what percent among the cases
    HCVData1 %>% filter(HCC_2019 == 1) %>% summarise(n(), mean(ImputedFIB4_2015), mean(ImputedFIB4_2018))

#Because Y is HCC occurence, Y2 is HCV diagnoses in 2016...
summary(HCVData$HCC_2016)
summary(HCVData$HCC_2017)
summary(HCVData$HCC_2018)
summary(HCVData$HCC_2019)


#So that our w's will represent the distributions in the population

racep1 = race[1]/nrow(HCVData1)
racep2 = race[2]/nrow(HCVData1)
racep3 = race[3]/nrow(HCVData1)
racep4 = race[4]/nrow(HCVData1)
sexp1 = sex[1]/nrow(HCVData1)
sexp2 = sex[2]/nrow(HCVData1)
SESp1 = SES[1]/nrow(HCVData1)
SESp2 = SES[2]/nrow(HCVData1)

w1 = racep1 * sexp1 * SESp1
w2 = racep1 * sexp1 * SESp2
w3 = racep1 * sexp2 * SESp1
w4 = racep1 * sexp2 * SESp1

w5 = racep2 * sexp1 * SESp1
w6 = racep2 * sexp1 * SESp2
w7 = racep2 * sexp2 * SESp1
w8 = racep2 * sexp2 * SESp2

w9 = racep3 * sexp1 * SESp1
w10 = racep3 * sexp1 * SESp2
w11 = racep3 * sexp2 * SESp1
w12 = racep3 * sexp2 * SESp2

w13 = racep4 * sexp1 * SESp1
w14 = racep4 * sexp1 * SESp2
w15 = racep4 * sexp2 * SESp1
w16 = racep4 * sexp2 * SESp2
```

```{r Data Structure 3 Simulation From R Lab 1, cache = TRUE}
#2. write function to generate O = (L(1), A(1), Y(2), ...Y(3))
generate_data  <- function(n) {
  U <- data.frame(UW = runif(n = n, min = 0, max = 1),
                  ULa1 = rnbinom(n=n, size = 4, prob = 0.6), 
                  ULa2 = rnbinom(n=n, size = 4, prob = 0.6), 
                  ULa3 = rnbinom(n=n, size = 4, prob = 0.6), 
                  ULa4 = rnbinom(n=n, size = 4, prob = 0.6), 
                  ULa5 = rnbinom(n=n, size = 4, prob = 0.6),  
                  ULb1 = rgamma(n = n, shape = 2, scale = 1), 
                  ULb2 = rgamma(n = n, shape = 2, scale = 1), 
                  ULb3 = rgamma(n = n, shape = 2, scale = 1), 
                  ULb4 = rgamma(n = n, shape = 2, scale = 1), 
                  ULb5 = rgamma(n = n, shape = 2, scale = 1), 
                  UA1 = runif(n=n, min = 0, max = 1),
                  UA2 = runif(n=n, min = 0, max = 1),
                  UA3 = runif(n=n, min = 0, max = 1),
                  UA4 = runif(n=n, min = 0, max = 1),
                  UA5 = runif(n=n, min = 0, max = 1),
                  UY2 = runif(n=n, min = 0, max = 1),
                  UY3 = runif(n=n, min = 0, max = 1),
                  UY4 = runif(n=n, min = 0, max = 1),
                  UY5 = runif(n=n, min = 0, max = 1))
  
  O <- U
  O <- O %>% 
    mutate(W = ifelse(UW < w1, 1,
                      ifelse(UW<(w1+w2), 2,
                             ifelse(UW<(w1+w2+w3), 3, 
                                    ifelse(UW<(w1+w2+w3+w4),4,
                                           ifelse(UW<(w1+w2+w3+w4+w5), 5,
                                                  ifelse(UW<(w1+w2+w3+w4+w5+w6), 6, 
                                                         ifelse(UW<(w1+w2+w3+w4+w5+w6+w7), 7, 
                        ifelse(UW<(w1+w2+w3+w4+w5+w6+w7+w8), 8, 
                               ifelse(UW<(w1+w2+w3+w4+w5+w6+w7+w8+w9), 9,
                                      ifelse(UW<(w1+w2+w3+w4+w5+w6+w7+w8+w9+w10), 10,
                                             ifelse(UW<(w1+w2+w3+w4+w5+w6+w7+w8+w9+w10+w11), 11,
                        ifelse(UW<(w1+w2+w3+w4+w5+w6+w7+w8+w9+w10+w11+w12), 12,
                                ifelse(UW<(w1+w2+w3+w4+w5+w6+w7+w8+w9+w10+w11+w12+w13), 13,
                        ifelse(UW<(w1+w2+w3+w4+w5+w6+w7+w8+w9+w10+w11+w12+w13+w14), 14,
                                ifelse(UW<(w1+w2+w3+w4+w5+w6+w7+w8+w9+w10+w11+w12+w13+w14+w15), 15, 16))))))))))))))),
           La1 = ULa1,
           Lb1 = ULb1,
           A1 = as.numeric(UA1 < 0.33* plogis(La1 + Lb1)),
           Y2 = as.numeric(UY2 < 0.07*plogis(Lb1 - La1 - A1)),
           
           La2 = ifelse(Y2==0, La1 + 0.1*ULa2, NA),
           Lb2 = ifelse(Y2==0, as.integer(Lb1 + 0.1*ULb2), NA),
           A2  = ifelse(Y2==0, as.numeric(UA2 < 0.33* plogis((La2 + Lb2 + A1))),NA),
           Y3  = ifelse(Y2==0, as.numeric(UY3 < 0.07*plogis(Lb2 - La2 - A2)),1),
           
           La3 = ifelse(Y3==0, La2 + 0.1*ULa3, NA),
           Lb3 = ifelse(Y3==0, as.integer(Lb2 + 0.1*ULb3), NA),
           A3  = ifelse(Y3==0, as.numeric(UA3 < 0.33* plogis((La3 + Lb3 + A2))),NA),
           Y4  = ifelse(Y3==0, as.numeric(UY4 < 0.07*plogis(Lb3 - La3 - A3)),1),
           
           La4 = ifelse(Y4==0, La3 + 0.1*ULa4, NA),
           Lb4 = ifelse(Y4==0, as.integer(Lb3 + 0.1*ULb4), NA),
           A4  = ifelse(Y4==0, as.numeric(UA4 < 0.33* plogis((La4 + Lb4 + A3))),NA),
           Y5  = ifelse(Y4==0, as.numeric(UY5 < 0.07*plogis(Lb4 - La4 - A4)),1)
           )
  return(O)
}

SimData <- generate_data(100000)
#skimr::skim(SimData)
```

## E. Estimation of our Target Causal Parameter

The target causal parameter is $\psi^F (\mathcal{P}_{U,X}) = E_{U,X}(Y_{\bar{a}=1(5)}-Y_{\bar{a}=0(5)})$, the average treatment effect on HCC diagnosis in year 5 assuming that all patients had abdominal imaging every year for the 4 prior years ($Y_{\bar{a}=1}$), compared to HCC outcomes in year 5 if all patients did NOT have abdominal imaging at any time point during the preceding four years ($Y_{\bar{a}=0}$). 

<!--Our target causal parameter is identified as a function of the observed data distribution under the assumptions that there is independence between each of the exogenous variables (i.e. there are no shared unknown variables influencing each of the endogenous nodes in our SCM) and that there are no practical positivity violations (i.e. there is a >0 probability of HCC among all of the baseline covariate and treatment regime combinations). In addition to controlling for any baseline covariates ($W$s), we also rely on the sequential randomization assumption for our data generating process.-->

```{r Intervene on Our Data structure, cache = TRUE}
# 3.3 - intervene on SCM/data generating function
generate_data_intervene  <- function(n, a) {
  U <- data.frame(UW = runif(n = n, min = 0, max = 1),
                  ULa1 = rnbinom(n=n, size = 4, prob = 0.6), 
                  ULa2 = rnbinom(n=n, size = 4, prob = 0.6), 
                  ULa3 = rnbinom(n=n, size = 4, prob = 0.6), 
                  ULa4 = rnbinom(n=n, size = 4, prob = 0.6), 
                  ULa5 = rnbinom(n=n, size = 4, prob = 0.6),  
                  ULb1 = rgamma(n = n, shape = 2, scale = 1), 
                  ULb2 = rgamma(n = n, shape = 2, scale = 1), 
                  ULb3 = rgamma(n = n, shape = 2, scale = 1), 
                  ULb4 = rgamma(n = n, shape = 2, scale = 1), 
                  ULb5 = rgamma(n = n, shape = 2, scale = 1), 
                  UY2 = runif(n=n, min = 0, max = 1),
                  UY3 = runif(n=n, min = 0, max = 1),
                  UY4 = runif(n=n, min = 0, max = 1),
                  UY5 = runif(n=n, min = 0, max = 1))
 
  O <- U
  O <- O %>%
    
 mutate(W = ifelse(UW < w1, 1,
                      ifelse(UW<(w1+w2), 2,
                             ifelse(UW<(w1+w2+w3), 3, 
                                    ifelse(UW<(w1+w2+w3+w4),4,
                                           ifelse(UW<(w1+w2+w3+w4+w5), 5,
                                                  ifelse(UW<(w1+w2+w3+w4+w5+w6), 6, 
                                                         ifelse(UW<(w1+w2+w3+w4+w5+w6+w7), 7, 
                        ifelse(UW<(w1+w2+w3+w4+w5+w6+w7+w8), 8, 
                               ifelse(UW<(w1+w2+w3+w4+w5+w6+w7+w8+w9), 9,
                                      ifelse(UW<(w1+w2+w3+w4+w5+w6+w7+w8+w9+w10), 10,
                                             ifelse(UW<(w1+w2+w3+w4+w5+w6+w7+w8+w9+w10+w11), 11,
                        ifelse(UW<(w1+w2+w3+w4+w5+w6+w7+w8+w9+w10+w11+w12), 12,
                                ifelse(UW<(w1+w2+w3+w4+w5+w6+w7+w8+w9+w10+w11+w12+w13), 13,
                        ifelse(UW<(w1+w2+w3+w4+w5+w6+w7+w8+w9+w10+w11+w12+w13+w14), 14,
                                ifelse(UW<(w1+w2+w3+w4+w5+w6+w7+w8+w9+w10+w11+w12+w13+w14+w15), 15, 16))))))))))))))),
           La1 = ULa1,
           Lb1 = ULb1,
           A1 = a,
           Y2 = as.numeric(UY2 < 0.07*plogis(Lb1 - La1 - A1)),
           
           La2 = ifelse(Y2==0, La1 + 0.1*ULa2, NA),
           Lb2 = ifelse(Y2==0, as.integer(Lb1 + 0.1*ULb2), NA),
           A2  = ifelse(Y2==0, a, NA),
           Y3  = ifelse(Y2==0, as.numeric(UY3 < 0.07*plogis(Lb2 - La2 - A2)),1),
           
           La3 = ifelse(Y3==0, La2 + 0.1*ULa3, NA),
           Lb3 = ifelse(Y3==0, as.integer(Lb2 + 0.1*ULb3), NA),
           A3  = ifelse(Y3==0, a, NA),
           Y4  = ifelse(Y3==0, as.numeric(UY4 < 0.07*plogis(Lb3 - La3 - A3)),1),
           
           La4 = ifelse(Y4==0, La3 + 0.1*ULa4, NA),
           Lb4 = ifelse(Y4==0, as.integer(Lb3 + 0.1*ULb4), NA),
           A4  = ifelse(Y4==0, a, NA),
           Y5  = ifelse(Y4==0, as.numeric(UY5 < 0.07*plogis(Lb4 - La4 - A4)),1)
           )
  return(O)
}

SimDataIntervene.0 <- generate_data_intervene(100000, 0)

SimDataIntervene.1 <- generate_data_intervene(100000, 1)


# 3.4 - evaluate Psi.F
True.Psi.F= mean(SimDataIntervene.1$Y5-SimDataIntervene.0$Y5, na.rm = TRUE)
True.Psi.F
```

```{r ltmle run on our simulation with SuperLearner, cache=TRUE}
#SL.library <- c("SL.glm", "SL.bayesglm", "SL.mean")
SL.library <- c("SL.glm", "SL.step", "SL.step.forward", "SL.earth", "SL.gam", "SL.nnet", "SL.mean", "SL.bayesglm", "SL.randomForest") ## armadillo - note I removed SL.step.interaction and SL.glm.interaction from Veronica's list to save computation time because I'm not very worried about interaction in this model, and when those wrappers were used on the real data the Coef was almost universally zero in every case.

# results_sim <- ltmle::ltmle(SimData,
#                      Anodes= c("A1", "A2", "A3", "A4"),
#                      Lnodes = c("La1", "Lb1", "La2", "Lb2", "La3", "Lb3", "La4", "Lb4"),
#                      Ynodes = c("Y2","Y3", "Y4", "Y5"),
#                      abar = list(c(1,1,1,1), c(0,0,0,0)),
#                      survivalOutcome = TRUE,
#                      SL.library = SL.library,
#                      attr(SL.library, "return.fit")
# )
# 
# results_sim.gcomp <- ltmle::ltmle(SimData,
#                      Anodes= c("A1", "A2", "A3", "A4"),
#                      Lnodes = c("La1", "Lb1", "La2", "Lb2", "La3", "Lb3", "La4", "Lb4"),
#                      Ynodes = c("Y2","Y3", "Y4", "Y5"),
#                      abar = list(c(1,1,1,1), c(0,0,0,0)),
#                      survivalOutcome = TRUE,
#                      gcomp = TRUE,
#                      SL.library = SL.library,
#                      attr(SL.library, "return.fit")
# )
# 
# save(results_sim, file = "results_sim.Rdata")
# save(results_sim.gcomp, file = "results_sim_gcomp.Rdata")


#loading results pre-run with simple SL library for knitting, etc.
load("results_sim.Rdata")
load("results_sim_gcomp.Rdata")

simtmle_SL <- summary(results_sim, "tmle")
simiptw_SL <- summary(results_sim, "iptw")
simgcomp_SL <- summary(results_sim.gcomp, "gcomp")

simtmle_ATE <- simtmle_SL$effect.measures$ATE$estimate; simtmle_ATE
simiptw_ATE <- simiptw_SL$effect.measures$ATE$estimate; simiptw_ATE
simgcomp_ATE <- simgcomp_SL$effect.measures$ATE$estimate; simgcomp_ATE

# performance of algorithms
results_sim$fit
results_sim.gcomp$fit

```

```{r Performance of estimators}

options(scipen = 999)

Bias.g <- simgcomp_ATE - True.Psi.F
Variance.g <- (simgcomp_SL$effect.measures$ATE$std.dev)^2
MSE.g <- Variance.g + Bias.g^2
rBias.g <- as.character(round(Bias.g, 6))
rVariance.g <- as.character(round(Variance.g, 7))
rMSE.g <- as.character(round(MSE.g, 7))

Bias.iptw <- simiptw_ATE - True.Psi.F
Variance.iptw <- (simiptw_SL$effect.measures$ATE$std.dev)^2
MSE.iptw <- Variance.iptw + Bias.iptw^2

Bias.tmle <- simtmle_ATE - True.Psi.F
Variance.tmle <- (simtmle_SL$effect.measures$ATE$std.dev)^2
MSE.tmle <- Variance.tmle + Bias.tmle^2
rVariance.tmle <- as.character(round(Variance.tmle, 5))
rMSE.tmle <- as.character(round(MSE.tmle, 5))
```

We have estimated our target causal parameter using G-computation, Inverse Probability of Treatment Weighting (IPTW), and LTMLE estimators within the \tt{ltmleMSM} \fontfamily{cmr}\selectfont function in the \textit{ltmle} R package. We implemented SuperLearner using generalized linear modeling (GLM), Bayesian GLM, generalized additive modeling (GAM), stepwise regression, stepwise regression with forward selection, adaptive regression splines, neural network analysis, a random forest algorithm, and mean of Y analyses to optimize our estimation.

When testing our estimators against a simulation of 100,000 patients roughly matching the variable distribution of our observed dataset, they performed with minimal bias and low overall mean squared error, with G-computation performing best overall (see Table below).

|| G-Comp | IPTW | TMLE |
|----------| ------ | ------ | ------ |
| Bias | `r rBias.g` | `r round(Bias.iptw, 6)` | `r round(Bias.tmle, 5)` |
| Variance | `r rVariance.g` | `r round(Variance.iptw, 6)` | `r rVariance.tmle` |
| MSE | `r rMSE.g` | `r round(MSE.iptw, 6)` | `r rMSE.tmle` |

The bias seen in the IPTW estimator is likely a result of practical positivity violations in our observed data. 

```{r Reconfigure Dataset, warning = FALSE, message = FALSE, cache = TRUE}

O1a <- as.data.frame(matrix(nrow=nrow(HCVData1), NA))
 
O1a$Y2 <- HCVData1$HCC_2016
O1a$Y3 <- HCVData1$HCC_2017
O1a$Y4 <- HCVData1$HCC_2018
O1a$Y5 <- HCVData1$HCC_2019

O1a$A1 <- HCVData1$ABD_IMAGING_YN_2015
O1a$A2 <- HCVData1$ABD_IMAGING_YN_2016
O1a$A3 <- HCVData1$ABD_IMAGING_YN_2017
O1a$A4 <- HCVData1$ABD_IMAGING_YN_2018

O1a$La1 <- ifelse(HCVData1$VISIT_COUNT_2015>0,1,0)
O1a$La2 <- ifelse(HCVData1$VISIT_COUNT_2016>0,1,0)
O1a$La3 <- ifelse(HCVData1$VISIT_COUNT_2017>0,1,0)
O1a$La4 <- ifelse(HCVData1$VISIT_COUNT_2018>0,1,0)

O1a$Lb1 <- HCVData1$FIB4_by2015
O1a$Lb2 <- HCVData1$FIB4_by2016
O1a$Lb3 <- HCVData1$FIB4_by2017
O1a$Lb4 <- HCVData1$FIB4_by2018

O1a$Lb1lm <- HCVData1$YearsSinceFib4_2015
O1a$Lb2lm <- HCVData1$YearsSinceFib4_2016
O1a$Lb3lm <- HCVData1$YearsSinceFib4_2017
O1a$Lb4lm <- HCVData1$YearsSinceFib4_2018

O1a$SES <- HCVData1$PayorType
O1a$Sex <- HCVData1$Sex
O1a$Race <- HCVData1$race.ethnicity
  
#O1a$C5 <- ifelse(O$A5 == 0, 1, 0)  armadillo
  
# #remove NAs before we imputed (now commented out) 
#   O1a <- subset(O, !is.na(O1a$SES))
#   O1a <- subset(O, !is.na(O1a$Race))
  
  O1a <- O1a %>% dplyr::select(SES, Race, Sex, La1, Lb1, Lb1lm, A1, 
                           Y2, La2, Lb2, Lb2lm, A2, 
                           Y3, La3, Lb3, Lb3lm, A3,
                           Y4, La4, Lb4, Lb4lm, A4, 
                           Y5)

O2a <- as.data.frame(matrix(nrow=nrow(HCVData2), NA))
 
O2a$Y2 <- HCVData2$HCC_2016
O2a$Y3 <- HCVData2$HCC_2017
O2a$Y4 <- HCVData2$HCC_2018
O2a$Y5 <- HCVData2$HCC_2019

O2a$A1 <- HCVData2$ABD_IMAGING_YN_2015
O2a$A2 <- HCVData2$ABD_IMAGING_YN_2016
O2a$A3 <- HCVData2$ABD_IMAGING_YN_2017
O2a$A4 <- HCVData2$ABD_IMAGING_YN_2018

O2a$La1 <- ifelse(HCVData2$VISIT_COUNT_2015>0,1,0)
O2a$La2 <- ifelse(HCVData2$VISIT_COUNT_2016>0,1,0)
O2a$La3 <- ifelse(HCVData2$VISIT_COUNT_2017>0,1,0)
O2a$La4 <- ifelse(HCVData2$VISIT_COUNT_2018>0,1,0)

O2a$Lb1 <- HCVData2$FIB4_by2015
O2a$Lb2 <- HCVData2$FIB4_by2016
O2a$Lb3 <- HCVData2$FIB4_by2017
O2a$Lb4 <- HCVData2$FIB4_by2018

O2a$Lb1lm <- HCVData2$YearsSinceFib4_2015
O2a$Lb2lm <- HCVData2$YearsSinceFib4_2016
O2a$Lb3lm <- HCVData2$YearsSinceFib4_2017
O2a$Lb4lm <- HCVData2$YearsSinceFib4_2018

O2a$SES <- HCVData2$PayorType
O2a$Sex <- HCVData2$Sex
O2a$Race <- HCVData2$race.ethnicity

  O2a <- O2a %>% dplyr::select(SES, Race, Sex, La1, Lb1, Lb1lm, A1, 
                           Y2, La2, Lb2, Lb2lm, A2, 
                           Y3, La3, Lb3, Lb3lm, A3,
                           Y4, La4, Lb4, Lb4lm, A4, 
                           Y5)

  
O3a <- as.data.frame(matrix(nrow=nrow(HCVData3), NA))
 
O3a$Y2 <- HCVData3$HCC_2016
O3a$Y3 <- HCVData3$HCC_2017
O3a$Y4 <- HCVData3$HCC_2018
O3a$Y5 <- HCVData3$HCC_2019

O3a$A1 <- HCVData3$ABD_IMAGING_YN_2015
O3a$A2 <- HCVData3$ABD_IMAGING_YN_2016
O3a$A3 <- HCVData3$ABD_IMAGING_YN_2017
O3a$A4 <- HCVData3$ABD_IMAGING_YN_2018

O3a$La1 <- ifelse(HCVData3$VISIT_COUNT_2015>0,1,0)
O3a$La2 <- ifelse(HCVData3$VISIT_COUNT_2016>0,1,0)
O3a$La3 <- ifelse(HCVData3$VISIT_COUNT_2017>0,1,0)
O3a$La4 <- ifelse(HCVData3$VISIT_COUNT_2018>0,1,0)

O3a$Lb1 <- HCVData3$FIB4_by2015
O3a$Lb2 <- HCVData3$FIB4_by2016
O3a$Lb3 <- HCVData3$FIB4_by2017
O3a$Lb4 <- HCVData3$FIB4_by2018

O3a$Lb1lm <- HCVData3$YearsSinceFib4_2015
O3a$Lb2lm <- HCVData3$YearsSinceFib4_2016
O3a$Lb3lm <- HCVData3$YearsSinceFib4_2017
O3a$Lb4lm <- HCVData3$YearsSinceFib4_2018

O3a$SES <- HCVData3$PayorType
O3a$Sex <- HCVData3$Sex
O3a$Race <- HCVData3$race.ethnicity

  O3a <- O3a %>% dplyr::select(SES, Race, Sex, La1, Lb1, Lb1lm, A1, 
                           Y2, La2, Lb2, Lb2lm, A2, 
                           Y3, La3, Lb3, Lb3lm, A3,
                           Y4, La4, Lb4, Lb4lm, A4, 
                           Y5)
  
O4a <- as.data.frame(matrix(nrow=nrow(HCVData4), NA))
 
O4a$Y2 <- HCVData4$HCC_2016
O4a$Y3 <- HCVData4$HCC_2017
O4a$Y4 <- HCVData4$HCC_2018
O4a$Y5 <- HCVData4$HCC_2019

O4a$A1 <- HCVData4$ABD_IMAGING_YN_2015
O4a$A2 <- HCVData4$ABD_IMAGING_YN_2016
O4a$A3 <- HCVData4$ABD_IMAGING_YN_2017
O4a$A4 <- HCVData4$ABD_IMAGING_YN_2018

O4a$La1 <- ifelse(HCVData4$VISIT_COUNT_2015>0,1,0)
O4a$La2 <- ifelse(HCVData4$VISIT_COUNT_2016>0,1,0)
O4a$La3 <- ifelse(HCVData4$VISIT_COUNT_2017>0,1,0)
O4a$La4 <- ifelse(HCVData4$VISIT_COUNT_2018>0,1,0)

O4a$Lb1 <- HCVData4$FIB4_by2015
O4a$Lb2 <- HCVData4$FIB4_by2016
O4a$Lb3 <- HCVData4$FIB4_by2017
O4a$Lb4 <- HCVData4$FIB4_by2018

O4a$Lb1lm <- HCVData4$YearsSinceFib4_2015
O4a$Lb2lm <- HCVData4$YearsSinceFib4_2016
O4a$Lb3lm <- HCVData4$YearsSinceFib4_2017
O4a$Lb4lm <- HCVData4$YearsSinceFib4_2018

O4a$SES <- HCVData4$PayorType
O4a$Sex <- HCVData4$Sex
O4a$Race <- HCVData4$race.ethnicity

  O4a <- O4a %>% dplyr::select(SES, Race, Sex, La1, Lb1, Lb1lm, A1, 
                           Y2, La2, Lb2, Lb2lm, A2, 
                           Y3, La3, Lb3, Lb3lm, A3,
                           Y4, La4, Lb4, Lb4lm, A4, 
                           Y5)

O5a <- as.data.frame(matrix(nrow=nrow(HCVData5), NA))
 
O5a$Y2 <- HCVData5$HCC_2016
O5a$Y3 <- HCVData5$HCC_2017
O5a$Y4 <- HCVData5$HCC_2018
O5a$Y5 <- HCVData5$HCC_2019

O5a$A1 <- HCVData5$ABD_IMAGING_YN_2015
O5a$A2 <- HCVData5$ABD_IMAGING_YN_2016
O5a$A3 <- HCVData5$ABD_IMAGING_YN_2017
O5a$A4 <- HCVData5$ABD_IMAGING_YN_2018

O5a$La1 <- ifelse(HCVData5$VISIT_COUNT_2015>0,1,0)
O5a$La2 <- ifelse(HCVData5$VISIT_COUNT_2016>0,1,0)
O5a$La3 <- ifelse(HCVData5$VISIT_COUNT_2017>0,1,0)
O5a$La4 <- ifelse(HCVData5$VISIT_COUNT_2018>0,1,0)

O5a$Lb1 <- HCVData5$FIB4_by2015
O5a$Lb2 <- HCVData5$FIB4_by2016
O5a$Lb3 <- HCVData5$FIB4_by2017
O5a$Lb4 <- HCVData5$FIB4_by2018

O5a$Lb1lm <- HCVData5$YearsSinceFib4_2015
O5a$Lb2lm <- HCVData5$YearsSinceFib4_2016
O5a$Lb3lm <- HCVData5$YearsSinceFib4_2017
O5a$Lb4lm <- HCVData5$YearsSinceFib4_2018

O5a$SES <- HCVData5$PayorType
O5a$Sex <- HCVData5$Sex
O5a$Race <- HCVData5$race.ethnicity

  O5a <- O5a %>% dplyr::select(SES, Race, Sex, La1, Lb1, Lb1lm, A1, 
                           Y2, La2, Lb2, Lb2lm, A2, 
                           Y3, La3, Lb3, Lb3lm, A3,
                           Y4, La4, Lb4, Lb4lm, A4, 
                           Y5)
```

```{r ltmle run on our DATA with SuperLearner, cache=TRUE}

# SL.library <- c("SL.glm", "SL.step", "SL.step.forward", "SL.earth", "SL.gam", "SL.nnet", "SL.mean", "SL.bayesglm", "SL.randomForest") 
# SL.library2 <- c("SL.glm", "SL.glm.interaction", "SL.step", "SL.step.forward", "SL.step.interaction", "SL.earth", "SL.gam", "SL.nnet", "SL.mean", "SL.bayesglm", "SL.randomForest") ## armadillo 
# 
# results1 <- ltmle::ltmle(O1a,
#                      Anodes= c("A1", "A2", "A3", "A4"),
#                      Lnodes = c("La1", "Lb1", "Lb1lm", "La2", "Lb2", "Lb2lm", "La3", "Lb3", "Lb3lm", "La4", "Lb4", "Lb4lm"),
#                      Ynodes = c("Y2","Y3", "Y4", "Y5"),
#                      abar = list(c(1,1,1,1), c(0,0,0,0)),
#                      survivalOutcome = TRUE,
#                      SL.library = SL.library2,
#                      attr(SL.library, "return.fit")
# )
# 
# results.gcomp1 <- ltmle::ltmle(O1a,
#                      Anodes= c("A1", "A2", "A3", "A4"),
#                      Lnodes = c("La1", "Lb1", "Lb1lm", "La2", "Lb2", "Lb2lm", "La3", "Lb3", "Lb3lm", "La4", "Lb4", "Lb4lm"),
#                      Ynodes = c("Y2","Y3", "Y4", "Y5"),
#                      abar = list(c(1,1,1,1), c(0,0,0,0)),
#                      survivalOutcome = TRUE,
#                      gcomp = TRUE,
#                      SL.library = SL.library2,
#                      attr(SL.library, "return.fit")
# )
# 
# 
# results2 <- ltmle::ltmle(O2a,
#                      Anodes= c("A1", "A2", "A3", "A4"),
#                      Lnodes = c("La1", "Lb1", "Lb1lm", "La2", "Lb2", "Lb2lm", "La3", "Lb3", "Lb3lm", "La4", "Lb4", "Lb4lm"),
#                      Ynodes = c("Y2","Y3", "Y4", "Y5"),
#                      abar = list(c(1,1,1,1), c(0,0,0,0)),
#                      survivalOutcome = TRUE,
#                      SL.library = SL.library2,
#                      attr(SL.library, "return.fit")
# )
# 
# results.gcomp2 <- ltmle::ltmle(O2a,
#                      Anodes= c("A1", "A2", "A3", "A4"),
#                      Lnodes = c("La1", "Lb1", "Lb1lm", "La2", "Lb2", "Lb2lm", "La3", "Lb3", "Lb3lm", "La4", "Lb4", "Lb4lm"),
#                      Ynodes = c("Y2","Y3", "Y4", "Y5"),
#                      abar = list(c(1,1,1,1), c(0,0,0,0)),
#                      survivalOutcome = TRUE,
#                      gcomp = TRUE,
#                      SL.library = SL.library2,
#                      attr(SL.library, "return.fit")
# )
# 
# 
# results3 <- ltmle::ltmle(O3a,
#                      Anodes= c("A1", "A2", "A3", "A4"),
#                      Lnodes = c("La1", "Lb1", "Lb1lm", "La2", "Lb2", "Lb2lm", "La3", "Lb3", "Lb3lm", "La4", "Lb4", "Lb4lm"),
#                      Ynodes = c("Y2","Y3", "Y4", "Y5"),
#                      abar = list(c(1,1,1,1), c(0,0,0,0)),
#                      survivalOutcome = TRUE,
#                      SL.library = SL.library2,
#                      attr(SL.library, "return.fit")
# )
# 
# results.gcomp3 <- ltmle::ltmle(O3a,
#                      Anodes= c("A1", "A2", "A3", "A4"),
#                      Lnodes = c("La1", "Lb1", "Lb1lm", "La2", "Lb2", "Lb2lm", "La3", "Lb3", "Lb3lm", "La4", "Lb4", "Lb4lm"),
#                      Ynodes = c("Y2","Y3", "Y4", "Y5"),
#                      abar = list(c(1,1,1,1), c(0,0,0,0)),
#                      survivalOutcome = TRUE,
#                      gcomp = TRUE,
#                      SL.library = SL.library2,
#                      attr(SL.library, "return.fit")
# )
# 
# 
# results4 <- ltmle::ltmle(O4a,
#                      Anodes= c("A1", "A2", "A3", "A4"),
#                      Lnodes = c("La1", "Lb1", "Lb1lm", "La2", "Lb2", "Lb2lm", "La3", "Lb3", "Lb3lm", "La4", "Lb4", "Lb4lm"),
#                      Ynodes = c("Y2","Y3", "Y4", "Y5"),
#                      abar = list(c(1,1,1,1), c(0,0,0,0)),
#                      survivalOutcome = TRUE,
#                      SL.library = SL.library2,
#                      attr(SL.library, "return.fit")
# )
# 
# results.gcomp4 <- ltmle::ltmle(O4a,
#                      Anodes= c("A1", "A2", "A3", "A4"),
#                      Lnodes = c("La1", "Lb1", "Lb1lm", "La2", "Lb2", "Lb2lm", "La3", "Lb3", "Lb3lm", "La4", "Lb4", "Lb4lm"),
#                      Ynodes = c("Y2","Y3", "Y4", "Y5"),
#                      abar = list(c(1,1,1,1), c(0,0,0,0)),
#                      survivalOutcome = TRUE,
#                      gcomp = TRUE,
#                      SL.library = SL.library2,
#                      attr(SL.library, "return.fit")
# )
# 
# 
# results5 <- ltmle::ltmle(O5a,
#                      Anodes= c("A1", "A2", "A3", "A4"),
#                      Lnodes = c("La1", "Lb1", "Lb1lm", "La2", "Lb2", "Lb2lm", "La3", "Lb3", "Lb3lm", "La4", "Lb4", "Lb4lm"),
#                      Ynodes = c("Y2","Y3", "Y4", "Y5"),
#                      abar = list(c(1,1,1,1), c(0,0,0,0)),
#                      survivalOutcome = TRUE,
#                      SL.library = SL.library2,
#                      attr(SL.library, "return.fit")
# )
# 
# results.gcomp5 <- ltmle::ltmle(O5a,
#                      Anodes= c("A1", "A2", "A3", "A4"),
#                      Lnodes = c("La1", "Lb1", "Lb1lm", "La2", "Lb2", "Lb2lm", "La3", "Lb3", "Lb3lm", "La4", "Lb4", "Lb4lm"),
#                      Ynodes = c("Y2","Y3", "Y4", "Y5"),
#                      abar = list(c(1,1,1,1), c(0,0,0,0)),
#                      survivalOutcome = TRUE,
#                      gcomp = TRUE,
#                      SL.library = SL.library2,
#                      attr(SL.library, "return.fit")
# )
# 
# save(results1, file = "results1.Rdata")
# save(results.gcomp1, file = "results_gcomp1.Rdata")
# save(results2, file = "results2.Rdata")
# save(results.gcomp2, file = "results_gcomp2.Rdata")
# save(results3, file = "results3.Rdata")
# save(results.gcomp3, file = "results_gcomp3.Rdata")
# save(results4, file = "results4.Rdata")
# save(results.gcomp4, file = "results_gcomp4.Rdata")
# save(results5, file = "results5.Rdata")
# save(results.gcomp5, file = "results_gcomp5.Rdata")
```

```{r results}

#loading results pre-run with SL2 library for knitting, etc.
load("results1.Rdata") 
load("results_gcomp1.Rdata")
load("results2.Rdata") 
load("results_gcomp2.Rdata") 
load("results3.Rdata") 
load("results_gcomp3.Rdata") 
load("results4.Rdata") 
load("results_gcomp4.Rdata") 
load("results5.Rdata") 
load("results_gcomp5.Rdata") 

tmle_SL1 <- summary(results1, "tmle")
iptw_SL1 <- summary(results1, "iptw")
gcomp_SL1 <- summary(results.gcomp1, "gcomp")
tmle_SL2 <- summary(results2, "tmle")
iptw_SL2 <- summary(results2, "iptw")
gcomp_SL2 <- summary(results.gcomp2, "gcomp")
tmle_SL3 <- summary(results3, "tmle")
iptw_SL3 <- summary(results3, "iptw")
gcomp_SL3 <- summary(results.gcomp3, "gcomp")
tmle_SL4 <- summary(results4, "tmle")
iptw_SL4 <- summary(results4, "iptw")
gcomp_SL4 <- summary(results.gcomp4, "gcomp")
tmle_SL5 <- summary(results5, "tmle")
iptw_SL5 <- summary(results5, "iptw")
gcomp_SL5 <- summary(results.gcomp5, "gcomp")

tmle_ATE1 <- exp(tmle_SL1$effect.measures$ATE$estimate); tmle_ATE1
iptw_ATE1 <- exp(iptw_SL1$effect.measures$ATE$estimate); iptw_ATE1
gcomp_ATE1 <- exp(gcomp_SL1$effect.measures$ATE$estimate); gcomp_ATE1
tmle_ATE2 <- exp(tmle_SL2$effect.measures$ATE$estimate); tmle_ATE2
iptw_ATE2 <- exp(iptw_SL2$effect.measures$ATE$estimate); iptw_ATE2
gcomp_ATE2 <- exp(gcomp_SL2$effect.measures$ATE$estimate); gcomp_ATE2
tmle_ATE3 <- exp(tmle_SL3$effect.measures$ATE$estimate); tmle_ATE3
iptw_ATE3 <- exp(iptw_SL3$effect.measures$ATE$estimate); iptw_ATE3
gcomp_ATE3 <- exp(gcomp_SL3$effect.measures$ATE$estimate); gcomp_ATE3
tmle_ATE4 <- exp(tmle_SL4$effect.measures$ATE$estimate); tmle_ATE4
iptw_ATE4 <- exp(iptw_SL4$effect.measures$ATE$estimate); iptw_ATE4
gcomp_ATE4 <- exp(gcomp_SL4$effect.measures$ATE$estimate); gcomp_ATE4
tmle_ATE5 <- exp(tmle_SL5$effect.measures$ATE$estimate); tmle_ATE5
iptw_ATE5 <- exp(iptw_SL5$effect.measures$ATE$estimate); iptw_ATE5
gcomp_ATE5 <- exp(gcomp_SL5$effect.measures$ATE$estimate); gcomp_ATE5
tmle_ATE <- as.vector(c(tmle_ATE1, tmle_ATE2, tmle_ATE3, tmle_ATE4, tmle_ATE5))
iptw_ATE <- as.vector(c(iptw_ATE1, iptw_ATE2, iptw_ATE3, iptw_ATE4, iptw_ATE5))
gcomp_ATE <- as.vector(c(gcomp_ATE1, gcomp_ATE2, gcomp_ATE3, gcomp_ATE4, gcomp_ATE5))

# average results across imputed datasets 
tmle_est_df <- as.data.frame(rbind(tmle_SL1$effect.measures$ATE$estimate, tmle_SL2$effect.measures$ATE$estimate, tmle_SL3$effect.measures$ATE$estimate, tmle_SL4$effect.measures$ATE$estimate, tmle_SL5$effect.measures$ATE$estimate))
tmle_SE_df <- as.data.frame(rbind(tmle_SL1$effect.measures$ATE$std.dev, tmle_SL2$effect.measures$ATE$std.dev, tmle_SL3$effect.measures$ATE$std.dev, tmle_SL4$effect.measures$ATE$std.dev, tmle_SL5$effect.measures$ATE$std.dev))
tmle_comb <- mi.meld(tmle_est_df, tmle_SE_df, byrow = TRUE)
tmle_ATE <- mean(tmle_ATE)
tmle_low <- round(tmle_ATE-1.96*exp(tmle_comb$se.mi),3)
tmle_high <- round(tmle_ATE+1.96*exp(tmle_comb$se.mi), 3)
iptw_est_df <- as.data.frame(rbind(iptw_SL1$effect.measures$ATE$estimate, iptw_SL2$effect.measures$ATE$estimate, iptw_SL3$effect.measures$ATE$estimate, iptw_SL4$effect.measures$ATE$estimate, iptw_SL5$effect.measures$ATE$estimate))
iptw_SE_df <- as.data.frame(rbind(iptw_SL1$effect.measures$ATE$std.dev, iptw_SL2$effect.measures$ATE$std.dev, iptw_SL3$effect.measures$ATE$std.dev, iptw_SL4$effect.measures$ATE$std.dev, iptw_SL5$effect.measures$ATE$std.dev))
iptw_comb <- mi.meld(iptw_est_df, iptw_SE_df, byrow = TRUE)
iptw_ATE <- mean(iptw_ATE)
iptw_low <- round(iptw_ATE-1.96*exp(iptw_comb$se.mi),3)
iptw_high <- round(iptw_ATE+1.96*exp(iptw_comb$se.mi), 3)
gcomp_est_df <- as.data.frame(rbind(gcomp_SL1$effect.measures$ATE$estimate, gcomp_SL2$effect.measures$ATE$estimate, gcomp_SL3$effect.measures$ATE$estimate, gcomp_SL4$effect.measures$ATE$estimate, gcomp_SL5$effect.measures$ATE$estimate))
gcomp_SE_df <- as.data.frame(rbind(gcomp_SL1$effect.measures$ATE$std.dev, gcomp_SL2$effect.measures$ATE$std.dev, gcomp_SL3$effect.measures$ATE$std.dev, gcomp_SL4$effect.measures$ATE$std.dev, gcomp_SL5$effect.measures$ATE$std.dev))
gcomp_comb <- mi.meld(gcomp_est_df, gcomp_SE_df, byrow = TRUE)
gcomp_ATE <- mean(gcomp_ATE)
gcomp_low <- round(gcomp_ATE-1.96*exp(gcomp_comb$se.mi), 3)
gcomp_high <- round(gcomp_ATE+1.96*exp(gcomp_comb$se.mi),3)

# performance of algorithms
results1$fit
results.gcomp1$fit
results2$fit
results.gcomp2$fit
results3$fit
results.gcomp3$fit
results4$fit
results.gcomp4$fit
results5$fit
results.gcomp5$fit

#NNT
NNTdenom <- gcomp_ATE/100
NNT <- round(1/(gcomp_ATE/100), 0)
```

## F. Findings

Ultimately, implementing \textit{ltmle} with \textit{Super Learner} on our observed data produced the following estimates for the risk difference of developing HCC in year 5 when receiving abdominal imaging at least once per year for the 4 preceding years, compared to not receiving abdominal imaging during that time, when controlling for the number of hepatology or primary care visits, FIB-4 score, sex, race, and SES was:

\vspace{12pt}

| **G-Comp** (95% CI) | **IPTW** (95% CI) | **TMLE** (95% CI) |
| ------------- | ------------- | ------------- |
| `r round(gcomp_ATE, 3)` (`r gcomp_low`, `r gcomp_high`) | `r round(iptw_ATE, 3)` (`r iptw_low`, `r iptw_high`) | `r round(tmle_ATE, 3)` (`r tmle_low`, `r tmle_high`) |

\newpage

## G. Interpretation

Using any of our estimators, we found that over the five years under study, patients with no cirrhosis at baseline were `r round(gcomp_ATE, 1)`% less likely to develop HCC by year 5 when they had abdominal imaging each year for the 4 previous years, compared to patients who had no abdominal imaging, when controlling for liver cirrhosis, number of primary care or hepatology visits, sex, race, and SES. While a `r round(gcomp_ATE, 1)`% decrease in HCC risk over 5 years sounds non-significant at first glance, this as a clinically meaningful finding given that we excluded patients with evidence of cirrhosis at baseline and only followed patients for 5 years. **This represents a number needed to treat (NNT) of `r NNT`**: for every `r NNT` people who receive annual abdominal imaging despite no evidence of cirrhosis at baseline, 1 case of HCC could be prevented over only 5 years. Given that abdominal imaging is a non-invasive and relatively inexpensive intervention, this underscores the potential impact of an intervention to educate primary care and hepatology providers about the importance of ensuring their chronically HCV-infected patients receive abdominal imaging at least once per year, regardless of liver cirrhosis.

## H. References

Axley P, Ahmed Z, Ravi S, Singal AK. Hepatitis C Virus and Hepatocellular Carcinoma: A Narrative Review. J Clin Transl Hepatol. 2018 Mar 28; 6(1): 79-84.

Bosetti C, Turati F, La Vecchia C. Hepatocellular carcinoma epidemiology.Best Pract Res Clin Gastroenterol. 2014 Oct; 28(5):753-70.

<!--
Burman BE, Bacchetti P, Khalili M. Moderate Alcohol Use and Insulin Action in Chronic Hepatitis C Infection. Dig Dis Sci. 2016;61(8):2417-2425. doi:10.1007/s10620-016-4119-0-->

Burstow NJ, Mohamed Z, Gomaa AI, et al. Hepatitis C treatment: where are we now? Int J Gen Med 2017; 10:39-52.

Calle EE, Rodriguez C, Walker-Thurmond K, Thun MJ. Overweight, obesity, and mortality from cancer in a prospectively studied cohort of U.S. adults. N Engl J Med. 2003 Apr 24; 348(17):1625-38.

Chuang SC, Lee YC, Hashibe M, Dai M, Zheng T, Boffetta P. Interaction between cigarette smoking and hepatitis B and C virus infection on the risk of liver cancer: a meta-analysis. Cancer Epidemiol Biomarkers Prev. 2010 May; 19(5):1261-8.

Donato F, Tagger A, Gelatti U, Parrinello G, Boffetta P, Albertini A, Decarli A, Trevisi P, Ribero ML, Martelli C, Porru S, Nardi G. Alcohol and hepatocellular carcinoma: the effect of lifetime intake and hepatitis virus infections in men and women. Am J Epidemiol. 2002 Feb 15; 155(4):323-31.

El-Serag HB. Epidemiology of viral hepatitis and hepatocellular carcinoma. Gastroenterology. 2012 May; 142(6):1264-1273.e1.

European Association for Study of Liver (EASL). EASL Recommendations on Treatment of Hepatitis C 2015. 
J Hepatol. 2015 Jul; 63(1):199-236.

Huang TS, Lin CL, Lu MJ, Yeh CT, Liang KH, Sun CC, Shyu YC, Chien RN. Diabetes, hepatocellular carcinoma, and mortality in hepatitis C-infected patients: A population-based cohort study.
J Gastroenterol Hepatol. 2017 Jul; 32(7):1355-1362.

Khan MQ, Anand V, Hessefort N, Hassan A, Ahsan A, Sonnenberg A, Fimmel CJ. Utility of Electronic Medical record-based Fibrosis Scores in Predicting Advanced Cirrhosis in Patients with Hepatitic C Virus Infection. J Transl Int Med. 2017 Mar; 5(1): 43-48.

Omland LH, Krarup H, Jepsen P, Georgsen J, Harritsh?j LH, Riisom K, Jacobsen SE, Schouenborg P, Christensen PB, S?rensen HT, Obel N, DANVIR Cohort Study. Mortality in patients with chronic and cleared hepatitis C viral infection: a nationwide cohort study. J Hepatol. 2010 Jul; 53(1):36-42.

Sterling, R. K., Lissen, E. , Clumeck, N. , Sola, R. , Correa, M. C., Montaner, J. , S. Sulkowski, M. , Torriani, F. J., Dieterich, D. T., Thomas, D. L., Messinger, D. and Nelson, M. (2006), Development of a simple noninvasive index to predict significant fibrosis in patients with HIV/HCV coinfection. Hepatology, 43: 1317-1325. doi:10.1002/hep.21178