---
title: "forSteph_tohelp"
author: "Shelley Facente"
date: "8/7/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

```{r data}
load("HIV_care.Rdata")

#This file was created using the code below

#HIV_care <- dataset %>% 
#  group_by(Acct.Number) %>% select(HIV_Care, Visit.Date.Time) %>%
#  arrange(Acct.Number, Visit.Date.Time)
```

The question is, how can I change HIV_care$HIV_Care so that when a single person (Acct.Number) has HIV_Care=="yes" AFTER HIV_Care=="no" (i.e. on a later visit they are in care, despite not being in care earlier) it recodes to "Linked" (i.e. linked to care), and similarly if they are "yes" but then at a later visit change to "no" it sets HIV_Care to "Lost"?

So in the end I would expect "Yes" (always in care), "No" (never in care), "Linked" (started out of care, then entered care during the timeframe of the dataset), and "Lost" (started in care, then fell out of care during the timeframe of the dataset).

Please help??? I'm sure there is dplyr magic but I don't know it. Blech.

```{r Hard-coded up to 4 visits-commented out}
##This Version only worked for datasets up to 4 visits

# HIV_Care <- HIV_Care %>%
#                 group_by(Acct.Number) %>%
#                 mutate(VisitNumber = 1:n()) %>% #just makes a visit number index bc using dates as column names is messy
#                 pivot_wider(names_from = VisitNumber, values_from = c(HIV_Care, Visit.Date.Time)) #makes one row per subject
# 
# #The combo of mutate and case_when allows you to create variables, based on conditions contained in the other variables. Coding always and never is easy, as you can see I've coded lost and linked based on the only 3 variations that can exist for that, based on only having four visits. There may be some general way of doing this for a higher number of visits automatically but I'd have to think more about that. 
# 
# HIV_Care <- HIV_Care %>%
#                 mutate( CareStatus = case_when(
#                   (HIV_Care_1 == "Yes" | is.na(HIV_Care_1) ==TRUE) &  
#                     (HIV_Care_2 == "Yes" | is.na(HIV_Care_2) ==TRUE) & 
#                       (HIV_Care_3 == "Yes" | is.na(HIV_Care_3) ==TRUE) &
#                         (HIV_Care_4 == "Yes" | is.na(HIV_Care_4) ==TRUE) ~ "Always",
#                   (HIV_Care_1 == "No" | is.na(HIV_Care_1) ==TRUE) &  
#                     (HIV_Care_2 == "No" | is.na(HIV_Care_2) ==TRUE) & 
#                       (HIV_Care_3 == "No" | is.na(HIV_Care_3) ==TRUE) &
#                         (HIV_Care_4 == "No" | is.na(HIV_Care_4) ==TRUE) ~ "Never",
#                   (HIV_Care_1 == "Yes") &  
#                     (HIV_Care_2 == "No" | is.na(HIV_Care_2) ==TRUE) | 
#                       (HIV_Care_3 == "No" | is.na(HIV_Care_3) ==TRUE) |
#                         (HIV_Care_4 == "No" | is.na(HIV_Care_4) ==TRUE) ~ "Lost",
#                   (HIV_Care_2 == "Yes") &  
#                       (HIV_Care_3 == "No" | is.na(HIV_Care_3) ==TRUE) |
#                         (HIV_Care_4 == "No" | is.na(HIV_Care_4) ==TRUE) ~ "Lost",
#                   (HIV_Care_3 == "Yes") &  
#                         (HIV_Care_4 == "No" | is.na(HIV_Care_4) ==TRUE) ~ "Lost",
#                   (HIV_Care_1 == "No") &  
#                     (HIV_Care_2 == "Yes" | is.na(HIV_Care_2) ==TRUE) | 
#                       (HIV_Care_3 == "Yes" | is.na(HIV_Care_3) ==TRUE) |
#                         (HIV_Care_4 == "Yes" | is.na(HIV_Care_4) ==TRUE) ~ "Linked",
#                   (HIV_Care_2 == "No") &  
#                       (HIV_Care_3 == "Yes" | is.na(HIV_Care_3) ==TRUE) |
#                         (HIV_Care_4 == "Yes" | is.na(HIV_Care_4) ==TRUE) ~ "Linked",
#                   (HIV_Care_3 == "No") &  
#                         (HIV_Care_4 == "Yes" | is.na(HIV_Care_4) ==TRUE) ~ "Linked"))

```

```{r Generalizable regardless of visit number}

#This solution is generalizable regardless of the number of visits

HIV_Care <- HIV_Care %>%
                 group_by(Acct.Number) %>% #so that R is doing manipulations within a single "account"'s set pf data
                 mutate(VisitNumber = 1:n())  #just makes a visit number index 

LastNotNAVisit <- HIV_Care %>%
            filter(!is.na(HIV_Care)) %>% #filter out NA rows
            mutate(LastNotNAVisit = case_when( VisitNumber == max(VisitNumber) ~ "Last")) %>% 
            #Marks the visit with the highest visit number, by creating a variable with that value, only when those conditions are met.
            filter(LastNotNAVisit == "Last") %>% #filters to only the last visit rows
            select(Acct.Number, VisitNumber, LastNotNAVisit) #makes just a little helper dataset to mark which is the last not NA visit

Care_Status <- left_join(HIV_Care, LastNotNAVisit) #left join means it keeps all the rows in HIV_Care, so basically we've just added a column to HIV_Care that says whether a particular row is the last non-NA visit or not.

rm(LastNotNAVisit) #just getting rid of this helper dataset

Care_Status <- Care_Status %>%
                 mutate(CareStatus = case_when( #again this creates a variable, and assigns it values based on other variables
                      LastNotNAVisit == "Last" & HIV_Care == "Yes" & all(HIV_Care == "Yes", na.rm = TRUE) ~ "Always",
                      LastNotNAVisit == "Last" & HIV_Care == "Yes" & any(HIV_Care == "No", na.rm = TRUE) ~ "Linked",
                      LastNotNAVisit == "Last" & HIV_Care == "No" & all(HIV_Care == "No", na.rm = TRUE) ~ "Never",
                      LastNotNAVisit == "Last" & HIV_Care == "No" & any(HIV_Care == "Yes", na.rm = TRUE) ~ "Lost")) %>%
                  select(Acct.Number, CareStatus) %>%
                  filter(!is.na(CareStatus))

#This leaves you with a minimal dataset that has acct number and care status, that you can merge back in to whatever you'd like. eg.
HIV_Care <- left_join(HIV_Care, Care_Status)

```
